---
title: "Serum_IgA_2022"
author: "Shruti Kakan"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading Libraries ####
```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
#library(PCAtools)
library(tidyverse)
library(Biobase)
#library(marray)
library(limma)
library(gplots)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
#devtools::install_github('renozao/NMF@devel')
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
#library(discreteMTP)
library(scales)
library(ggsci)
library(ggplot2)

```

# IgA Analysis ##########
#______________________________________________________________

```{r Reading in Raw IgA Serum Data}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/')
IgA_NSI <- read.csv("IgA_data/IgA_MCF_SSK_546_NSI_norm_Serum.csv",  header=T)[1:80,]
IgA_SNR <- read.csv("IgA_data/IgA_MCF_SSK_546_SNR_Serum.csv",  header=T)[1:80,]
Strain <- c(rep("NOD", each=5),rep("NOR", each=3), rep("BALBc", each=6))

```

#Adding Auto-antigen id names 
```{r Adding Auto-antigen id names }
#Adding Auto-antigen id names 
Antigen_ID <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/Antigen_ID.csv",  header=T)[1:80,1:2]

#colnames(IgA_NSI)[1] <- colnames(Antigen_ID)[2]
#colnames(IgA_SNR)[1] <- colnames(Antigen_ID)[2]

Antigen_ID[68,"ID"] <- IgA_NSI[68,"ID"]
IgA_NSI <- full_join(IgA_NSI, Antigen_ID, by="ID")
IgA_SNR <- full_join(IgA_SNR, Antigen_ID, by="ID")
rownames(IgA_NSI) <- IgA_NSI$Antigen_ID
rownames(IgA_SNR) <- IgA_SNR$Antigen_ID

IgA_NSI <- IgA_NSI[,-c(1, 17)]
IgA_SNR <- IgA_SNR[,-c(1, 17)]

```


```{r}
####Setting up column metadata

Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))

colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain))
colnames(colData) <- c('Sample', "Strain")
rownames(colData) <- colData$Sample
colData$Strain <- factor(colData$Strain)
#colData$Strain <- relevel(colData$Strain, ref = "BALBc")
#Biofluid <- c( rep("Tear", each=11), rep("Serum", each=11))
#colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain, Biofluid))
```

#### Filtering Data based on low Signal to Noise ratio
```{r Filtering rows with SNR<3}
IgA_raw=list()
IgA_SNR$average <- rowMeans(IgA_SNR[,-1])
IgA_SNR$med <- rowMedians(as.matrix(IgA_SNR[-1]))
IgA_NSI$average <- rowMeans(IgA_NSI[,-1])
IgA_NSI$med <- rowMedians(as.matrix(IgA_NSI[-1]))

IgA_raw$NSI <- as.matrix(IgA_NSI[which(IgA_NSI$med>3 & IgA_SNR$med>3),])[,1:15]
IgA_raw$SNR <- as.matrix(IgA_SNR[which(IgA_NSI$med>3 & IgA_SNR$med>3),])[,1:15]
```

#### Visualizing Filtered Data ####
```{r}

dataN <- list()
dataN[["Raw"]]<- log(IgA_NSI[,2:15] * IgA_SNR[,2:15] + 0.5) 
dataN[["filtered"]] <- log(IgA_raw$NSI * IgA_raw$SNR + 0.5)[,-1] 

for (i in 1:2) {
   countData = as.data.frame(dataN[[i]])
   df_dseq = melt(countData, variable.name = "Samples", value.name = "count")# reshape the matrix 
   mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(15)
   p <- ggplot(df_dseq, aes(x = count, color=Samples)) +
      geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
      theme_minimal() + #xlim(-1.5,6) +
      scale_colour_manual(values=mycolors, name="") +
      guides(fill="none")
   plot(p)
}


```

NOD M1 Serum appears to be the most variable, and was therefore removed

#### Boxplots of raw and SNR based Filtered Data ####

```{r}
boxplot(as.data.frame((dataN$filtered)),main="IgA normalization")#,col=Sample)

dataN <- (IgA_NSI[,1:15] * IgA_SNR[,1:15] ) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI*SNR")

dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +1) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Log2 NSI*SNR")

dataN <- log2(IgA_raw$NSI + 1) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="log2 NSI")
```

#IgA levels appear to be much lower in serum than in tears, especially in male NOD mice


#Limma Based Quantile Normalization
```{r}
library(limma)
dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5)[,2:15]
mydata <- as.matrix(dataN)

conditions<- paste(colData$Strain[-1],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design, plot=TRUE)

plot.new()

cont.matrix<- makeContrasts(
   NTvBT =  NOD - BALBc,
   nTvBT = NOR - BALBc,
   levels = design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont, robust=TRUE)
plotMDS(fit, col=as.numeric(conditions))
plotMA(fit)
plotMA(fit.cont, coef=1)
plotMA(fit.cont, coef=2)
#plot.new()
plotMD(fit.cont, col=2)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)

```

#### Voom Normalization After Removing outliers ####

NOD M1 was removed recursively
```{r Voom After removing outlier samples}

###### Voom normalization with quantiles
conditions<- paste(colData$Strain[-1],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

v <- voom(counts=(IgA_raw$NSI*IgA_raw$SNR + 1)[,-1], design, plot=TRUE, normalize="cyclicloess")

#v <- voom(dae, design, plot=TRUE)
fit <- lmFit(v, design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont, robust=TRUE)
topTable(fit.cont, number=20)
#topTable(fit, coef=ncol(design))

NTvsBT <- topTable(fit.cont, coef=1, number=30, adjust.method = 'BH')
nTvsBT <- topTable(fit.cont, coef=2, number=30, adjust.method = 'BH')
NTvsBT$Antigen <- row.names(NTvsBT)
nTvsBT$Antigen <- row.names(nTvsBT)
NTvnTvBT <- full_join(NTvsBT, nTvsBT, by="Antigen")
rownames(NTvnTvBT) <- NTvnTvBT$Antigen
NTvnTvBT[is.na(NTvnTvBT)] <- 0.5
NTvnTvBT <- NTvnTvBT[c(which(NTvnTvBT$adj.P.Val.x < 0.1 & NTvnTvBT$adj.P.Val.y < 0.1)),]
write.csv(NTvnTvBT, file="NTvnTvBT_Serum_IgA.csv", sep=',')
v$E["PM Scl100",]

plotMDS(v,col=as.numeric(Strain))
plot.new()
plotMA(fit)
plotMD(fit.cont, col=2)
plotMD(fit.cont, col=1)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)



```

#### Density plots post voom normalization ####
```{r Density plots}
dae <- DGEList(counts=(IgA_raw$NSI * IgA_raw$SNR + 0.5)[,-1])
keep <- filterByExpr(dae, design)
dae <- dae[keep,,keep.lib.sizes=FALSE]
dae <- calcNormFactors(dae)
logCPM <- cpm(dae, log=TRUE, prior.count=2)

df_dseq = melt(v$E, variable.name = "Samples", value.name = "count")# reshape the matrix 
#df_dseq = melt(logCPM, variable.name = "Samples", value.name = "count")# reshape the matrix 
df_dseq$Strain <- factor(substr(df_dseq$Var2, 1,3))
mycolors <- colorRampPalette(brewer.pal(8,"Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Var2)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + xlim(-5, 25) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")
```

##### R-Squared ....... goodness of fit ####
```{r}
for (i in 1:38){
   sst <- rowSums(v$E^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
plot(1:38, Rsq)
which(Rsq<0.90)

summary(fit.cont$r.squared)
```

###Determining DE genes
```{r}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```

#### DE Analysis using voom normalized counts
```{r}
results<-decideTests(fit.cont,adjust.method="none", p=0.05)
summary(results)

write.fit(fit.cont,file="IgA_Serum_DGE.csv",adjust="none",results=results, sep=',')
topTable(fit.cont)
fitObj <- read.csv("IgA_Serum_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]
write.csv(fitObjDE,file="IgA_Serum_Voom_Hits.csv", sep=',')
fitObjDE$Autoantigen <- rownames(fitObjDE)

mydata <- as.matrix(v$E)
```

#### Boxplots from voom normalized counts ####

```{r Boxplots for Serum IgA DE Genes}
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=10), hjust=0.4),,
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=16, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=10), face='bold', hjust=-0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

hits <- rownames(fitObjDE)
Y=matrix(nrow=18,ncol=14)
for (i in 1:18) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample[-1]
Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain[-1]
Y$Sample <- paste0(Y$Strain, c(2:5,1:4, 1:6))

for (i in 1:18){
   setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
   filename <- paste(hits[i],"IgA_Serum.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.6, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
      #scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
      ylab("Log Norm Intensity")  + 
      labs(title=colnames(Y)[i], hjust=0.5) +
      scale_fill_jco()
   tiff(filename, units="in", width=3.8, height=3.7, res=300)
   print(p)
   dev.off()
}
```

#### Combined Boxplot####

```{r}
Y_combined <- Y[,c(NTvnTvBT$Antigen[which(NTvnTvBT$adj.P.Val.x<0.05 & NTvnTvBT$adj.P.Val.y < 0.05 & NTvnTvBT$logFC.x>1.5 & NTvnTvBT$logFC.y>1.5)])]

Y_combined[,(ncol(Y_combined)+1):(ncol(Y_combined)+2)] <- Y[,(ncol(Y)-1):ncol(Y)]
Y_combined <- gather(Y_combined, "Antigen", "V Counts", 1:(ncol(Y_combined)-2))

chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=18, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=10), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

tiff("Serum_IgG_hits.tiff", units="in", width=5.5, height=4, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022")
ggplot(Y_combined, aes(x=Strain, y=`V Counts`, fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.65, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.5, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
   facet_wrap(~Antigen, ncol=2) +
      theme_minimal() +
      chart_design +
      ylab("Log Normalized Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
dev.off()
```

####IgG Analysis Serum #####
```{r}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

```

####Reading in Raw IgG Datasets ####
```{r}
setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/')
IgG_NSI <- read.csv("IgG_MCF_SSK_546_NSI_serum.csv",  header=T)[1:80,]
IgG_SNR <- read.csv("IgG_MCF_SSK_546_serum_SNR.csv",  header=T)[1:80,]
Strain <- c(rep("NOD", each=5),rep("NOR", each=3), rep("BALBc", each=6))
```

#####Autoantigen ID names ####
```{r Adding Auto-antigen id names }
#Adding Auto-antigen id names 
Antigen_ID <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/Antigen_ID.csv",  header=T)[1:80,1:2]

#colnames(IgG_NSI)[1] <- colnames(Antigen_ID)[2]
#colnames(IgG_SNR)[1] <- colnames(Antigen_ID)[2]

Antigen_ID[68,"ID"] <- IgG_NSI[68,"ID"]
IgG_NSI <- full_join(IgG_NSI, Antigen_ID, by="ID")
IgG_SNR <- full_join(IgG_SNR, Antigen_ID, by="ID")
rownames(IgG_NSI) <- IgG_NSI$Antigen_ID
rownames(IgG_SNR) <- IgG_SNR$Antigen_ID

IgG_NSI <- IgG_NSI[,-c(1, 17)]
IgG_SNR <- IgG_SNR[,-c(1, 17)]
```

#####Setting up column metadata ###
```{r}
Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))

colData <- as.data.frame(cbind(c(colnames(IgG_NSI)), Strain))
colnames(colData) <- c('Sample', "Strain")
rownames(colData) <- colData$Sample
colData$Strain <- factor(colData$Strain)
#colData$Strain <- relevel(colData$Strain, ref = "BALBc")
#Biofluid <- c( rep("Tear", each=11), rep("Serum", each=11))
#colData <- as.data.frame(cbind(c(colnames(IgG_NSI)), Strain, Biofluid))
```

#####Filtering Data based on low Signal to Noise ratio
```{r Filtering rows with SNR<3}
IgG_raw=list()
IgG_SNR$average <- rowMeans(IgG_SNR)
IgG_SNR$med <- rowMedians(as.matrix(IgG_SNR))

IgG_NSI$average <- rowMeans(IgG_NSI)
IgG_NSI$med <- rowMedians(as.matrix(IgG_NSI))

IgG_raw$NSI <- as.matrix(IgG_NSI[which(IgG_NSI$med>50 & IgG_SNR$med>2),])[,1:15]
IgG_raw$SNR <- as.matrix(IgG_SNR[which(IgG_NSI$med>50 & IgG_SNR$med>2),])[,1:15]
IgG_raw$Ab_Score <- as.matrix(log2((IgG_raw$NSI*IgG_raw$NSI) + 1))

```

#####Visualizing Filtered Data ####
```{r}
dataN <- log(IgG_raw$NSI * IgG_raw$SNR + 1) 
countData = as.data.frame(dataN)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count")# reshape the matrix 

mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Samples)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + #xlim(-1.5,6) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")
```

There does not appear to be any extreme outliers

#####Boxplots of raw and SNR based Filtered Data ####

```{r}
boxplot(as.data.frame((dataN)),main="IgG normalization")#,col=Sample)

dataN <- (IgG_raw$Ab_Score) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI*SNR")

dataN <- log(cpm(IgG_raw$NSI * IgG_raw$SNR)+0.5)
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Log2 NSI*SNR")

dataN <- log(IgG_raw$NSI*1000/colSums(IgG_raw$NSI) + 1.05) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI")

dataN <- (IgG_raw$Ab_Score*1000)/colSums(IgG_raw$Ab_Score) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Intensity Normalized NSI")
```
We see global differences between groups, with BALB/c medians being lower than both NOD and NOR.

#####Limma Based Normalization##
```{r}
conditions<- paste(colData$Strain,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

library(limma)
dataN <- log((IgG_raw$NSI*IgG_raw$SNR*1000000)/colSums(IgG_raw$NSI*IgG_raw$SNR)+1)
plotMD(dataN) #
fit <- lmFit(dataN, design, plot=TRUE)
plotMD(fit, column=1)

cont.matrix<- makeContrasts(
   NTvBT =  NOD - BALBc,
   nTvBT = NOR - BALBc,
   levels = design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont, robust=TRUE)

topTable(fit.cont, number=30)

NTvsBT <- topTable(fit.cont, coef=1, number=36, adjust.method = 'BH')
nTvsBT <- topTable(fit.cont, coef=2, number=36, adjust.method = 'BH')
NTvsBT$Antigen <- row.names(NTvsBT)
nTvsBT$Antigen <- row.names(nTvsBT)
NTvnTvBT <- full_join(NTvsBT, nTvsBT, by="Antigen")
rownames(NTvnTvBT) <- NTvnTvBT$Antigen
NTvnTvBT[is.na(NTvnTvBT)] <- 0.5
   NTvnTvBT <- NTvnTvBT[c(which(NTvnTvBT$adj.P.Val.x < 0.05 & NTvnTvBT$adj.P.Val.y < 0.05)),]

write.csv(NTvnTvBT, file="NTvnTvBT_Serum_IgG.csv", sep=',')

```

QC Plots

```{r}
plotMD(fit)
plotMD(fit.cont)
plotMDS(fit, col=as.numeric(Strain))
plotMDS(fit)
plotMA(fit)
plotMA(fit.cont, coef=1)
plotMA(fit.cont, coef=2)
#plot.new()
qqnorm(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)

```


Quantile Normalization is over-normalizing this dataset. Therefore we will do a simple log2 Normalization followed by limma based DE analysis.

Library Normalization appears to be the best recourse
#####Voom Normalization After Removing outliers ####

```{r Voom After removing outlier samples}

###### Voom normalization with quantiles
conditions<- paste(colData$Strain[-c(1)],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)
block <- as.factor(c(rep("N",9), rep("B",6)))
v <- voom(counts=exp(dataN[,-c(1)]), design, plot=TRUE, normalize="cyclicloess")

#v <- voom(dae, design, plot=TRUE)
fit <- lmFit(v, design, robust="TRUE")
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
topTable(fit.cont, number=40)
#topTable(fit, coef=ncol(design))


plotMDS(v,col=as.numeric(Strain[-c(1)]))
plot.new()
plotMA(fit.cont, coef=1)
plotMD(fit.cont, col=2)
plotMD(fit.cont, col=1)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)

```



#####Density plots post voom normalization ####
```{r Density plots}
df_dseq = melt(dataN, variable.name = "Samples", value.name = "count")# reshape the matrix 
df_dseq$Strain <- factor(substr(df_dseq$Var2, 1,3))
mycolors <- colorRampPalette(brewer.pal(8,"Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Var2)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + xlim(-5, 25) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")
```

#####R-Squared ....... goodness of fit ####
```{r}
for (i in 1:77){
   sst <- rowSums(v$E^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
plot(1:77, Rsq)
which(Rsq<0.90)

summary(fit.cont$r.squared)
```

#####Determining DE genes##
```{r}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```

#####DE Analysis using voom normalized counts
```{r}
results<-decideTests(fit.cont,adjust.method="fdr", p=0.05)
summary(results)

write.fit(fit.cont,file="IgG_Serum_DGE.csv",adjust="none",results=results, sep=',')
topTable(fit.cont, coef=1, 25)
topTable(fit.cont, coef=2, 25)
fitObj <- read.csv("IgG_Serum_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]
write.csv(fitObjDE,file="IgG_Serum_Limma_Hits.csv", sep=',')
fitObjDE$Autoantigen <- rownames(fitObjDE)
```

##### Boxplots from voom normalized counts ####

```{r Boxplots for Serum IgG DE Genes}
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=10), hjust=0.4),,
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=16, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=10), face='bold', hjust=-0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

mydata <- as.matrix(dataN)
#mydata <- as.matrix(v$E)
hits <- rownames(NTvnTvBT)
Y=matrix(nrow=12,ncol=15)
for (i in 1:12) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample
Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Sample <- paste0(Y$Strain, c(1:5,1:4, 1:6))

for (i in 1:12){
   setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
   filename <- paste(hits[i],"IgG_Serum.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
      #scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
      ylab("Log Norm Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco() #+ ylim(0,14)
   tiff(filename, units="in", width=3.8, height=4.7, res=300)
   print(p)
   dev.off()
}
```

#####Combined Boxplots ####
```{r}
Y_combined <- Y[,c(NTvnTvBT$Antigen[which(NTvnTvBT$adj.P.Val.x<0.05 & NTvnTvBT$adj.P.Val.y < 0.05 & NTvnTvBT$logFC.x>1.75 & NTvnTvBT$logFC.y>1.75)])]

Y_combined[,(ncol(Y_combined)+1):(ncol(Y_combined)+2)] <- Y[,(ncol(Y)-1):ncol(Y)]
Y_combined <- gather(Y_combined, "Antigen", "V Counts", 1:(ncol(Y_combined)-2))


chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),
   axis.text.x = element_text(size=13),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=17, margin = margin(r = 5)),
   strip.text.x = element_text(size =15, margin = margin(b=10), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

tiff("Serum_IgG_hits.tiff", units="in", width=8.5, height=6.5, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
ggplot(Y_combined, aes(x=Strain, y=`V Counts`, fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.65, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.5, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
   facet_wrap(~Antigen, ncol=4, scales="free_x") +
      theme_minimal() +
      chart_design +
      ylab("Log Normalized Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
dev.off()
```

