---
title: "Supplemental Methods 1"
subtitle: "Study 3 - Tear IgA Analysis"
author: "Shruti Kakan"
date: "Last Updated 2024-10-24"
output: html_document
---
### 1. Loading Libraries ####
```{r Setup, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = "~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/"
)
library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
#library(PCAtools)
library(tidyverse)
library(Biobase)
#library(marray)
library(limma)
library(gplots)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
#devtools::install_github('renozao/NMF@devel')
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
#library(discreteMTP)
library(scales)
library(ggsci)
library(ggplot2)

```

### ______________________________________________________________

### 2. Reading in Raw Tear Data

```{r Reading in Raw Tear Data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/')
#IgA_NSI <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_NSI_norm.csv",  header=T)[1:80,]
IgA_NSI <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_NSI.csv",  header=T)[1:80,]
IgA_SNR <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_SNR.csv",  header=T)[1:80,]
Strain <- c(rep("NOD", each=5),rep("NOR", each=3), rep("BALBc", each=6))
```

#### 2.1 Adding Auto-antigen Id names 
```{r Adding Auto-antigen id names, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Adding Auto-antigen id names 
Antigen_ID <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/Antigen_ID.csv",  header=T)[1:80,1:2]

#colnames(IgA_NSI)[1] <- colnames(Antigen_ID)[2]
#colnames(IgA_SNR)[1] <- colnames(Antigen_ID)[2]

Antigen_ID[68,"ID"] <- IgA_NSI[68,"ID"]
IgA_NSI <- full_join(IgA_NSI, Antigen_ID, by="ID")
IgA_SNR <- full_join(IgA_SNR, Antigen_ID, by="ID")
rownames(IgA_NSI) <- IgA_NSI$Antigen_ID
rownames(IgA_SNR) <- IgA_SNR$Antigen_ID

IgA_NSI <- IgA_NSI[,-c(1, 17)]
IgA_SNR <- IgA_SNR[,-c(1, 17)]

```

#### 2.2 Setting up column Metadata
```{r Column Metadata, message=FALSE, warning=FALSE}
Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))

colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain))
colnames(colData) <- c('Sample', "Strain")
rownames(colData) <- colData$Sample
colData$Strain <- factor(colData$Strain)
#colData$Strain <- relevel(colData$Strain, ref = "BALBc")
#Biofluid <- c( rep("Tear", each=11), rep("Serum", each=11))

#colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain, Biofluid))

```


### 3. Filtering Data based on low Signal to Noise ratio

Rows with low signal to noise ratio of less than rowmeans 2.8 were removed. Additionally rows with low overall signal intensity were also removed.
```{r Filtering rows with SNR<3, message=FALSE, warning=FALSE}
IgA_raw=list()
IgA_SNR$average <- rowMeans(as.matrix(IgA_SNR))
IgA_SNR$med <- rowMedians(as.matrix(IgA_SNR))

IgA_raw$NSI <- as.matrix(IgA_NSI[which(IgA_SNR$average>2.8),])
IgA_raw$SNR <- as.matrix(IgA_SNR[which(IgA_SNR$average>2.8),])[,1:15]
#IgA_raw$NSI <- as.matrix(IgA_NSI[which(rowSums(IgA_SNR[,1:15]>2.8) > 6),])
#IgA_raw$SNR <- as.matrix(IgA_SNR[which(rowSums(IgA_SNR[,1:15]>2.8) > 6),][,1:15])

log2(colSums(IgA_raw$NSI+0.5))

```
This reduced the total number of Autoantibodies (rows) included in the analysis from 80 to 69.

#### 3.1 Visualizing Filtered Data ####
```{r Desnity plots of raw data, message=FALSE, warning=FALSE, paged.print=FALSE}
dataN <- log2(IgA_NSI + 0.5) 
countData = as.data.frame(dataN)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count")# reshape the matrix 

mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Samples)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + #xlim(-1.5,6) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")
```
BALB/c Sample M6 appears to have extremely low signal intensity values across the board. 

```{r Boxplots of raw & filtered data, message=FALSE, warning=FALSE, paged.print=FALSE}
boxplot(as.data.frame((dataN)),main="Tear IgA Signal Intensity Prior to Row filtering")#,col=Sample)

#dataN <- (IgA_NSI * IgA_SNR[,1:15] ) 
dataN <- (IgA_NSI) 
countData = as.data.frame(dataN)
#boxplot(as.data.frame((dataN)),main="NSI")

#dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5) 
dataN <- log2(IgA_raw$NSI+0.5) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Tear IgA Signal Intensity After Row filtering")

```

### 4. Limma Based Default Normalization & DE Analysis

Sample BALB/c M6 was excluded from normalization and differential expression analysis.
```{r}
library(limma)
#dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5)[,-15]
dataN <- log2(IgA_raw$NSI +0.5)[,-15] #Sample BALB/c M6 removed
mydata <- as.matrix(dataN)

conditions<- paste(colData$Strain[-15],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)

cont.matrix<- makeContrasts(
   NTvBT =  NOD - BALBc,
   nTvBT = NOR - BALBc,
   levels = design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
```

#### 4.1 Default Limma Normalization QC
```{r Limma QC, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#MA plots showing relationship between average expression of autoantibodies and their log fold-changes
plot.new()
plotMD(fit.cont, col=1) # NOD v BALB/c
plotMD(fit.cont, col=2) # NOR v BALB/c

# QQ Plots of normalized data
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
#Density plots of normalized Data
plotDensities(fit)
```


While this normalization works fairly well, most of the BALB/c samples show a relatively low signal intensity across the board. While this my be a true biological effect, we have to consider it as a potential batch effect owing to the fact that BALB/c is a closely matched control of NOD and NOR.  

```{r Limma Default DE Analysis, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

#Generating results data frames using Limma's topTable() function
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
NTvsBT <- topTable(fit.cont, coef=1, p.value=1, number=50, adjust.method = 'BH')
nTvsBT <- topTable(fit.cont, coef=2, p.value=1, number=50, adjust.method = 'BH')
NTvsBT$Antigen <- row.names(NTvsBT)
nTvsBT$Antigen <- row.names(nTvsBT)
NTvnTvBT <- full_join(NTvsBT, nTvsBT, 
                      by="Antigen",
                      suffix = c(".NOD", ".NOR"))
rownames(NTvnTvBT) <- NTvnTvBT$Antigen
```

#### 4.2 Default Limma Results Table

```{r Limma Default - Results Table, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

#Generating output file
NTvnTvBT[is.na(NTvnTvBT)] <- 0.5
NTvnTvBT <- NTvnTvBT[-c(which(NTvnTvBT$adj.P.Val.NOD > 0.1 & NTvnTvBT$adj.P.Val.NOR > 0.1)),]
write.csv(NTvnTvBT, file="Tear_IgA_Limma.csv", sep=',')

print(NTvnTvBT[which(NTvnTvBT$logFC.NOD>1),c(1:2,5,8,12)])
```
16 IgA autoantibodies are significantly upregulated in both NOD & NOR as compared to male BALB/c with adjusted p value < 0.05

### 5. Voom Normalization After Removing outliers

Using the Limma Voom package and a quantile normalization approach, we hope to completely remove the batch effect. This should give us a more conservative result.

```{r Limma-Voom, Quantile Normalization, message=FALSE, warning=FALSE, paged.print=FALSE}
#After removing outlier samples BALB/c 6

###### Voom normalization with quantiles
conditions<- paste(colData$Strain[-15],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

v <- voom(counts=(IgA_raw$NSI[,-15]+0.5), design, plot=TRUE, normalize="quantile")
#v <- voom(dae, design, plot=TRUE)
fit <- lmFit(v, design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
#plot(x=colSums(IgA_raw$NSI), y=colSums(IgG_raw$NSI))
```

#### 5.1 QC plots of analysis
```{r}
#PC1 separates Diseases and healthy samples
plotMDS(v,col=as.numeric(Strain))

#MA plots showing relationship between average expression of autoantibodies and their log fold-changes
plot.new()
plotMD(fit.cont, col=1) # NOD v BALB/c
plotMD(fit.cont, col=2) # NOR v BALB/c

# The Q-Q plot is a fairly straight line
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)

# The Density plots of all three groups overlap fairly well, confirming that the batch effect has been largely normalized
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)
```

#### 5.2 Density plots of normalized data
```{r Density plots of Normalized Data by Strain, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Density plots

df_dseq = melt(v$E, variable.name = "Samples", value.name = "count")# reshape the matrix 
df_dseq$Strain <- factor(substr(df_dseq$Var2, 1,3))
mycolors <- colorRampPalette(brewer.pal(8,"Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Var2)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   facet_wrap(~Strain, ncol=2) +
   theme_minimal() + xlim(-5, 25) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")

```

Sample wise Density plots facetted by the mouse strain show that voom normalized Log2 transformed NSI data of each sample overlap quite well and follow normal distribution. 

#### 5.3 QC: Assessing Goodness of fit
```{r goodness of fit}
#R-Squared ....... goodness of fit
for (i in 1:69){
   sst <- rowSums(v$E^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
plot(1:69, Rsq)
which(Rsq<0.90)

summary(fit.cont$r.squared)
```

#### 5.4 Results Table: Limma-Voom (Quantile) Normalized DE 
```{r Limma-Voom Results Table, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
NTvsBT <- topTable(fit.cont, coef=1, p.value=1, number=50, adjust.method = 'BH')
nTvsBT <- topTable(fit.cont, coef=2, p.value=1, number=50, adjust.method = 'BH')
NTvsBT$Antigen <- row.names(NTvsBT)
nTvsBT$Antigen <- row.names(nTvsBT)
NTvnTvBT <- full_join(NTvsBT, nTvsBT, 
                      by="Antigen", 
                      suffix = c(".NOD", ".NOR"))
rownames(NTvnTvBT) <- NTvnTvBT$Antigen
NTvnTvBT[is.na(NTvnTvBT)] <- 0.5
NTvnTvBT <- NTvnTvBT[-c(which(NTvnTvBT$adj.P.Val.NOD > 0.1 & NTvnTvBT$adj.P.Val.NOR > 0.1)),]
write.csv(NTvnTvBT, file="Tear_IgA_voom.csv", sep=',')
print(NTvnTvBT[which(NTvnTvBT$logFC.NOD>1),c(1:2,5,8,12)])
```

Thirteen Auto-antibodies are significantly upregulated in tears of both male NOD & male NOR as compared to tears of healthy male BALB/c.  

### 6. Plotting Differentially Expressed Autoantibodies

```{r Comparing different methods of p-value correction, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```
#### 6.1 Print output for DE Analysis done using voom normalized counts 
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
results<-decideTests(fit.cont,adjust.method="none", p=0.005)
summary(results)

write.fit(fit.cont,file="IgA_Tears_DGE.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("IgA_Tears_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]
write.csv(fitObjDE,file="IgA_Tears_Voom_Hits.csv", sep=',')
fitObjDE$Autoantigen <- rownames(fitObjDE)

```

#### 6.2 Boxplots from voom normalized counts ####

```{r Individual Boxplots for Tear IgA DE Genes, message=FALSE, warning=FALSE, paged.print=FALSE}
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=18, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=20), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

mydata <- as.matrix(v$E)
hits <- rownames(NTvnTvBT)
hits[(length(hits)+1)] <- "Ro SSA 52"
#hits[(length(hits)+1)] <- "RoSSA 60"

Y=matrix(nrow=length(hits),ncol=14)

for (i in 1:length(hits)) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample[-15]

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain[-15]
Y$Sample <- paste0(Y$Strain, c(1:5,1:4, 1:5))

setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
for (i in length(hits)){
   filename <- paste(hits[i],"IgA_Tears.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
      scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
      ylab("Log Norm Intensity")  + 
      scale_x_discrete(labels= c("BALB/c", "NOD", "NOR")) + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
   tiff(filename, units="in", width=3.45, height=3.5, res=300)
   print(p)
   dev.off()
}
#print(p)
```

#### 6.3 Combined Boxplots of Upregulated hits 
```{r Combined Boxplots, echo=TRUE, fig.height=8.5, fig.width=12.2, message=FALSE, warning=FALSE, paged.print=FALSE}
chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),,
   axis.text.x = element_text(size=16, margin = margin(b=5)),
   axis.text.y = element_text(size=18),
   axis.title.x = element_blank(),
   legend.text = element_blank(),
   legend.title = element_blank(),
   legend.position = "top",
   axis.title.y = element_text(size=24, margin = margin(r = 5)),
   strip.text.x = element_text(size =19, margin = margin(b=15), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

#Generating Data Frame for plotting Combined boxplot
Y_combined <- Y[,c(NTvnTvBT$Antigen[which(NTvnTvBT$adj.P.Val.NOD<0.051 & NTvnTvBT$adj.P.Val.NOR < 0.051 & NTvnTvBT$logFC.NOD>0)])]
#Y_combined[,14:15] <- Y[,c("IF", "La SSB")]
Y_combined[,(ncol(Y_combined)+1):(ncol(Y_combined)+2)] <- Y[,(ncol(Y)-1):(ncol(Y))]
Y_combined <- gather(Y_combined, "Antigen", "V Counts", 1:(ncol(Y_combined)-2))
#Y_combined[which(Y_combined$Strain=="BALBc"),"Strain"] = factor("BALB/c")


#Generating combined boxplot of 13 autoantibodies using ggplot()
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
tiff("Tear_IgA_hits.tiff", units="in", width=12.2, height=8.5, res=300)
q <- ggplot(Y_combined, aes(x=Strain, y=`V Counts`, fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.7, coef=1, varwidth=F, show.legend = F, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.6, size=3.3, show.legend = F, position = position_jitterdodge(dodge.width=1.2))+
   facet_wrap(~Antigen, ncol=5, scales="free_x") +
      theme_minimal() +
      chart_design +
      ylab("Log Normalized Intensity")  + 
   scale_x_discrete(labels= c("BALB/c", "NOD", "NOR")) +
      labs(title=factor(Y_combined$Antigen), hjust=0.5) +
      scale_fill_jco()
print(q)
dev.off()

(q)
```


```{r}
Tear_IgAhits <- read.csv("IgA_Tears_Voom_Hits.csv", header=T)
Tear_IgGhits <- read.csv("IgG_Tears_Voom_Hits.csv", header=T)

Tear_Ig <- full_join(Tear_IgAhits[,c(1,5,6,7,8)], Tear_IgGhits[,c(1,5,6,7,8)], by="X")

write.csv(Tear_Ig,file="IgA_IgG_Overlap_Tears_Hits.csv", sep=',')
```


### 7. Code for Comparing Total Tear IgG and IgA Signal Intensity (ColSums) ##################

```{r}
Tear_IgG_NSI_raw <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgG_MCF_SSK_546_Tear_NSI.csv", header=TRUE, row.names = 1, nrows=88)
Tear_IgG_NSI_raw <- Tear_IgG_NSI_raw[,1:15]
Tear_IgG_SNR_raw <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgG_MCF_SSK_546_Tear_SNR.csv", header=TRUE, row.names = 1, , nrows=88)

Tear_IgA_NSI_raw <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/IgA_MCF_SSK_546_Tear_NSI.csv", header=TRUE, row.names = 1)
Tear_IgA_NSI_raw <- Tear_IgA_NSI_raw[1:88,]

x = colSums(Tear_IgG_NSI_raw[1:80,1:14])
y = colSums(Tear_IgA_NSI_raw[1:80,1:14])

data <- as.data.frame(cbind(x,y))
colnames(data) <- c("IgG","IgA")
Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=5))
data$Strain <- factor(Strain)
data$Sample <- rownames(data)
data <- gather(data, "Ig", "Intensity", 1:2)
```

 
#### 7.1 Total tear IgG total IgA intensity for 80 antigens (column sum) 

```{r Boxplots ColSum_IgA_vs_IgG.tiff, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),,
   axis.text.x = element_text(size=16, margin = margin(b=5)),
   axis.text.y = element_text(size=18),
   axis.title.x = element_blank(),
   legend.text = element_text(size=10),
   legend.title = element_blank(),
   legend.position = "top",
   axis.title.y = element_text(size=20, margin = margin(r = 5)),
   strip.text.x = element_text(size =19, margin = margin(b=15), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

p <- ggplot(data, aes(x=Strain, y=(Intensity), fill=Ig)) +
   geom_boxplot(outlier.shape = NA, width = 0.8, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
   geom_jitter(aes(colour = Ig), alpha =0.6, size=2.5, show.legend = T, position = position_jitterdodge(dodge.width=0.9)) +
   scale_fill_jco() + 
   theme_minimal() + 
   ylab("Total Intensity of Autoantibodies") +
   chart_design #+ 
   #ylim(1,(290000))

tiff("ColSum_IgA_vs_IgG.tiff", units="in", width=5.5, height=4.5, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
print(p)
dev.off()

print(p)
```


While sum of intensity of 80 autoantigens for IgA is higher in NOR mice tears than IgG, this increase is not statistically significant.

No significant difference was observed between the mean signal IgG intensity as compared to IgA signal intensity for the 80 autoantigens tested in either male NOD or BALB/c mice. 

#### 7.1.1 Statistics on ColSums(Normalized Intensity)
```{r Statistics for colSums, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

mod <- lm((Intensity) ~ Strain*Ig, data)
aov(mod)
summary(aov(mod))
TukeyHSD(aov(mod))
plot(mod, which = 2)
hist(mod$residuals)

shapiro.test(mod$residuals) #data are non-normal
plot(mod, which = 3)
#leveneTest(aov(Intensity ~ Strain + Ig, data=data))
```



#### 7.2 Chunk for plotting mouse IgG vs IgA secondary control for each samples

```{r Mouse IgG vs IgA control, message=FALSE, warning=FALSE, paged.print=FALSE}
#x1 = unlist(Tear_IgG_NSI_raw[87,2:16])
#x2 = unlist(Tear_IgG_NSI_raw[88,2:16])
#y1 = unlist(Tear_IgA_NSI_raw[87,])
#y2 = unlist(Tear_IgA_NSI_raw[88,])
Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))
x = unlist(lapply(Tear_IgG_NSI_raw[87:88,], mean, 2))
y = unlist(lapply(Tear_IgA_NSI_raw[87:88,], mean, 2))
data <- as.data.frame(cbind(x,y))
colnames(data) <- c("IgG","IgA")
data$Strain <- factor(Strain)
data$Sample <- rownames(data)
data <- gather(data, "Ig", "Intensity", 1:2)

tiff("Tear_IgA_vs_IgG.tiff", units="in", width=5.5, height=4.5, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
q <- ggplot(data, aes(x=Strain, y=(Intensity), fill=Ig)) +
   geom_boxplot(outlier.shape = NA, width = 0.8, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
   geom_jitter(aes(colour = Ig), alpha =0.6, size=2.5, show.legend = T, position = position_jitterdodge(dodge.width=0.9)) +
   scale_fill_jco() +
   theme_minimal() +
   ylab("Average Intensity of Ig Controls") +
   chart_design #+ 
   #ylim(0,100000)
print(q)
dev.off()

print(q)
   #+ ylim(10,12)
```

Signal intensity observed for mouse IgA secondary control was significantly higher than the intensity observed for mouse IgG secondary control in the tears of male NOD, NOR and BALB/c mice. Levels of signal intensity for either IgG or IgA were consistent between the three mouse groups.

#### 7.2.1 Statistics for mouse IgG and IgA controls

```{r Statistics for secondary controls, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Statistics 
mod <- lm((Intensity) ~ Strain*Ig, data)
aov(mod)
summary(aov(mod))
TukeyHSD(aov(mod))
plot(mod, which = 2)
hist(mod$residuals)

shapiro.test(mod$residuals) #data are normally distributed
plot(mod, which = 3)
```



### 8. Tear IgG vs IgA Correlation Analysis #########

Scatterplots showing raw signal intensity for IgG vs IgA of each autoantigen in the tears of NOD, NOR and BALB/c mice. 

```{r Correlation Analysis, echo=TRUE, fig.height=11.5, fig.width=16, message=FALSE, warning=FALSE, paged.print=FALSE}
#which((rowSums(Tear_IgG_SNR_raw[1:80,1:15])/15) < 2.8)
#which((rowSums(IgA_SNR[1:80,])/15) < 2.8)

Tear_IgA_NSI <- Tear_IgA_NSI_raw[1:80,]
Tear_IgA_NSI <- Tear_IgA_NSI[-c(which((rowSums(Tear_IgG_SNR_raw[1:80,1:15])/15) < 2.8 & (rowSums(IgA_SNR[1:80,])/15) < 2.8)),]
Tear_IgA_NSI$Antigen <- rownames(Tear_IgA_NSI)
Tear_IgA_NSI <- gather(Tear_IgA_NSI, "Sample", "Intensity", 1:(ncol(Tear_IgA_NSI)-1))
Tear_IgA_NSI$Strain <- str_sub(Tear_IgA_NSI$Sample, 1, (str_locate(Tear_IgA_NSI$Sample, "_")[,1] -1))
Tear_IgA_NSI$Ig <- "IgA"

Tear_IgG_NSI <- Tear_IgG_NSI_raw[1:80,]
Tear_IgG_NSI <- Tear_IgG_NSI[-c(which((rowSums(Tear_IgG_SNR_raw[1:80,])/15) < 2.8 & (rowSums(IgA_SNR[1:80,])/15) < 2.8)),]
Tear_IgG_NSI$Antigen <- rownames(Tear_IgG_NSI)
Tear_IgG_NSI <- gather(Tear_IgG_NSI, "Sample", "Intensity", 1:(ncol(Tear_IgG_NSI)-1))
Tear_IgG_NSI$Strain <- str_sub(Tear_IgG_NSI$Sample, 1, (str_locate(Tear_IgG_NSI$Sample, "_")[,1] -1))
Tear_IgG_NSI$Ig <- "IgG"

data_mat <- dplyr::bind_rows(Tear_IgG_NSI, Tear_IgA_NSI)

###Plot for NODs & NORs
data_mat2 <- dplyr::full_join(Tear_IgA_NSI, Tear_IgG_NSI, by = c("Antigen", "Sample"),suffix = c(".IgA", ".IgG"),)

### Plot for NORs
ggplot(data = data_mat2[which(data_mat2$Strain.IgA=="NOR"),]) +
   geom_point(aes(x=(Intensity.IgA), y=(Intensity.IgG), fill=factor(Antigen), color=factor(Antigen)), show.legend = F) +
   facet_wrap(~factor(Antigen)) 
```


There is a need to normalize the NSI on x and y axes. In the next plot, we run correlation on log10 normalized NSI for IgG and IgA.

#### 8.1 Correlation Scatterplots of log 10 IgG NSI vs log log IGA NSI
```{r Plot for BALB/c, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
### Plot for BALB/c
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022")
tiff("BALBc_IgGvsIgA_Tears.jpeg", units="in", width=16, height=9.5, res=300)
ggplot(data = data_mat2[which(data_mat2$Strain.IgA=="BALBc"),], aes(x=log10(Intensity.IgA+1), y=log10(Intensity.IgG+1))) +
   geom_point(aes( fill=factor(Antigen), color=factor(Antigen)), show.legend = F) +
   geom_smooth(method=lm, se=FALSE, fullrange=FALSE, linewidth=0.5) +
   stat_cor(color='black', method = "spearman", label.x=0, label.y = 4, p.digits=2, p.accuracy = 0.0001) +
   facet_wrap(~factor(Antigen),nrow=7) +
   chart_design
dev.off()

```


```{r DGCA analysis, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#### DGCA Analysis
BiocManager::install(c("GO.db", "AnnotationDbi", "impute", "preprocessCore"))
install.packages("WGCNA")
install.packages("devtools")
#For this github install to work you will need to update your personal access token or delete your personal access token.
#remotes::install_github("andymckenzie/DGCA")


#define design matrix as IgG vs IgA
n_IgA_samples = 15; n_IgG_samples = 15 
Ig_type = c(rep("IgA", n_oligo_samples), rep("IgG", n_neuron_samples))
design_mat = model.matrix(~ Ig_type + 0)
colnames(design_mat) = c("IgA", "IgG")
str(design_mat)
design_mat <- model.matrix(~0+ data_mat$Ig)

library(DGCA, quietly = TRUE)
ddcor_res = ddcorAll(inputMat = data_mat, design = design,
  compare = c("IgA", "IgG"),
  adjust = "none", heatmapPlot = FALSE, nPerm = 0, nPairs = 100)
head(ddcor_res)
```

#### 8.2 Correlation Analysis Figure
```{r Plot for NODs & NOR, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

chart_design <- theme(
   plot.title = element_text(color = "Black", size = 20, face = "bold", margin = margin(b=5), hjust=0.4),
   legend.position = 'bottom',
   legend.text = element_text(size=13, margin = margin(t=1, b=1)),
   axis.text.x = element_text(size=12),
   axis.text.y = element_text(size=12),
   axis.title.x = element_text(size=18, margin = margin(t=5, b=1)),
   axis.title.y = element_text(size=18, margin = margin(r=5)),
   strip.text.x = element_text(size =10, margin = margin(b=2), face='bold', hjust=0.4),
   strip.background = element_blank(), strip.placement = "outside")


p <- ggplot(data = data_mat2[which(data_mat2$Strain.IgA %in% c("NOD", "NOR")),], aes(x=log10(Intensity.IgA+1), y=log10(Intensity.IgG+1))) +
   geom_point(aes(fill=Strain.IgA, color=factor(Antigen)), pch=21, size=3, alpha = 0.8, show.legend = TRUE) +
   facet_wrap(~factor(Antigen), nrow=7) +
   geom_smooth(method=lm, se=FALSE, fullrange=FALSE, linewidth=0.5) +
   #scale_color_manual(values=col) +
   #scale_shape_manual(values=c(1:25,1:11)) +
   scale_fill_jco(guide="legend") +
   guides(color="none") +
   theme_minimal() +
   xlab("Log 10 IgA NSI") +
   ylab("Log 10 IgG NSI") +
   stat_cor(color='black', method = "spearman", label.x=0, label.y = 4, p.digits=2, p.accuracy = 0.0001) +
   chart_design

setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022")
tiff("ScatterPlot_IgGvsIgA_Tears.jpeg", units="in", width=16, height=9.5, res=300)
print(p)
dev.off()

print(p)
```

Sctterplots of IgG vs IgA for NOD, NOR and BALB/c.  Each point represents one mouse. Spearman correlation coefficient rho and p-values are shown. p < 0.01 is considered significant.

```{r Correlation Plot for NOD, echo=TRUE, fig.height=12.5, fig.width=16, message=FALSE, warning=FALSE, paged.print=FALSE}
### Plot for NOD, NOR & BALB/c
data_mat2$model <- ifelse(data_mat2$Strain.IgA=="BALBc", "Healthy", "SjD")

p <- ggplot(data = data_mat2, aes(x=log10(Intensity.IgA+1), y=log10(Intensity.IgG+1))) +
   geom_point(aes( color=factor(model), fill=factor(Antigen)), show.legend = F, alpha = 0.8) +
   geom_smooth(aes(color = factor(model)),method=lm, se=FALSE, fullrange=FALSE, linewidth=0.25) +
   scale_color_jco(guide="legend") +
   guides(fill="none") +
   theme_minimal() + ylim(-1, 4.6) +
   xlab("Log 10 IgA NSI") +
   ylab("Log 10 IgG NSI") +
   stat_cor(aes(color=factor(model)), method = "spearman", r.digits=2, p.accuracy=0.0005, label.y = c(-0.9,4.4), cor.coef.name="rho") +
   facet_wrap(~factor(Antigen),nrow=7) +
   chart_design

print(p)

setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022")
tiff("All_IgGvsIgA_Tears.jpeg", units="in", width=16, height=10, res=300)
print(p)
dev.off()
```

#### 8.3 Scatterplot of shared epitopes between upregulated IgG and IgA autoantibodies 

Scatterplot of Log10 IgG vs Log10 IgA NSI of shares epitopes of autoantibodies that are significantly upregulated in male NOD and male NOR mice tears when compared to BALB/c controls.
NOD and NOR are shown in yellow (referred to an SjD) and BALB/c are shown in blue (referred to as Healthy).

```{r Supplemental Figure 6, echo=TRUE, fig.height=4.8, fig.width=6.5, message=FALSE, warning=FALSE, paged.print=FALSE}

data_mat3 <- data_mat2[which(data_mat2$Antigen %in% c("tTG", "TPO", "Mi-2", "Jo-1", "IA-2", "SAE1/SAE2 ")),]
p <- ggplot(data = data_mat3, aes(x=log10(Intensity.IgA+1), y=log10(Intensity.IgG+1))) +
   geom_point(aes( color=factor(model), fill=factor(Antigen)), show.legend = T, alpha = 0.8) +
   geom_smooth(aes(color = factor(model)),method=lm, se=FALSE, fullrange=FALSE, linewidth=0.25) +
   scale_color_jco(guide="legend") +
   guides(fill="none") +
   theme_minimal() + ylim(-1, 4.6) +
   xlab("Log 10 IgA NSI") +
   ylab("Log 10 IgG NSI") +
   stat_cor(aes(color=factor(model)), method = "spearman", r.digits=2, p.accuracy=0.0005, label.y = c(-0.9,4.4), cor.coef.name="rho") +
   facet_wrap(~factor(Antigen),nrow=2) +
   chart_design

print(p)

setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022")
tiff("Supp_Figure_6.jpeg", units="in", width=6.5, height=4.8, res=300)
print(p)
dev.off()
```

