---
title: "NOD NOR Tear Serum"
author: "Shruti Kakan"
date: "2024-06-26"
output: html_document
---
#### Loading Libraries ####
```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
#library(PCAtools)
library(tidyverse)
library(Biobase)
#library(marray)
library(limma)
library(gplots)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
#devtools::install_github('renozao/NMF@devel')
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
#library(discreteMTP)
library(scales)
library(ggsci)
library(ggplot2)

```

# IgA Analysis ##########
#______________________________________________________________

```{r Reading in Raw Tear Data}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/')
IgA_NSI <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_NSI_norm.csv",  header=T)[1:80,]
IgA_SNR <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_SNR.csv",  header=T)[1:80,]
Strain <- c(rep("NOD", each=5),rep("NOR", each=3), rep("BALBc", each=6))

```


#Adding Auto-antigen id names 
```{r Adding Auto-antigen id names }
#Adding Auto-antigen id names 
Antigen_ID <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/Antigen_ID.csv",  header=T)[1:80,1:2]

#colnames(IgA_NSI)[1] <- colnames(Antigen_ID)[2]
#colnames(IgA_SNR)[1] <- colnames(Antigen_ID)[2]

Antigen_ID[68,"ID"] <- IgA_NSI[68,"ID"]
IgA_NSI <- full_join(IgA_NSI, Antigen_ID, by="ID")
IgA_SNR <- full_join(IgA_SNR, Antigen_ID, by="ID")
rownames(IgA_NSI) <- IgA_NSI$Antigen_ID
rownames(IgA_SNR) <- IgA_SNR$Antigen_ID

IgA_NSI <- IgA_NSI[,-c(1, 17)]
IgA_SNR <- IgA_SNR[,-c(1, 17)]

```


```{r}
####Setting up column metadata

Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))

colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain))
colnames(colData) <- c('Sample', "Strain")
rownames(colData) <- colData$Sample
colData$Strain <- factor(colData$Strain)
#colData$Strain <- relevel(colData$Strain, ref = "BALBc")
#Biofluid <- c( rep("Tear", each=11), rep("Serum", each=11))

#colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain, Biofluid))

```


#### Filtering Data based on low Signal to Noise ratio
```{r Filtering rows with SNR<3}
IgA_raw=list()
IgA_SNR$average <- rowMeans(IgA_SNR)
IgA_SNR$med <- rowMedians(as.matrix(IgA_SNR))

IgA_raw$NSI <- as.matrix(IgA_NSI[which(IgA_SNR$med>1.8),])
IgA_raw$SNR <- as.matrix(IgA_SNR[which(IgA_SNR$med>1.8),])[,1:15]

```

#### Visualizing Filtered Data ####
```{r}
dataN <- log2(IgA_NSI * IgA_SNR[,1:15] + 0.5) 
countData = as.data.frame(dataN)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count")# reshape the matrix 

mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Samples)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + #xlim(-1.5,6) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")



```


```{r}
boxplot(as.data.frame((dataN)),main="IgG normalization")#,col=Sample)

dataN <- (IgA_NSI * IgA_SNR[,1:15] ) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI*SNR")


dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Log2 NSI*SNR")


dataN <- log2(IgA_raw$NSI + 0.5) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI")

```

#Limma Based Quantile Normalization
```{r}
library(limma)
dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5) 
mydata <- as.matrix(dataN)

conditions<- paste(colData$Strain,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)

cont.matrix<- makeContrasts(
   NTvBT =  NOD - BALBc,
   nTvBT = NOR - BALBc,
   levels = design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)

plot.new()
plotMD(fit.cont, col=2)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)


```


#### Voom Normalization After Removing outliers
```{r}
#After removing outlier samples

###### Voom normalization with quantiles
conditions<- paste(colData$Strain,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

v <- voom(counts=(IgA_raw$NSI*IgA_raw$SNR+0.5), design, plot=TRUE, normalize="quantile")
#v <- voom(dae, design, plot=TRUE)
fit <- lmFit(v, design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
topTable(fit.cont)
#topTable(fit, coef=ncol(design))

plotMDS(v,col=as.numeric(Strain))
plot.new()
plotMD(fit.cont, col=2)
plotMD(fit.cont, col=1)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)

#Density plots

dae <- DGEList(counts=(IgA_raw$NSI*IgA_raw$SNR+0.5))
keep <- filterByExpr(dae, design)
dae <- dae[keep,,keep.lib.sizes=FALSE]
dae <- calcNormFactors(dae)
logCPM <- cpm(dae, log=TRUE, prior.count=2)

df_dseq = melt(v$E, variable.name = "Samples", value.name = "count")# reshape the matrix 
df_dseq$Strain <- factor(substr(df_dseq$Var2, 1,3))
mycolors <- colorRampPalette(brewer.pal(8,"Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Var2)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   facet_wrap(~Strain, ncol=2) +
   theme_minimal() + xlim(-5, 25) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")

#R-Squared ....... goodness of fit
for (i in 1:71){
   sst <- rowSums(v$E^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
plot(1:71, Rsq)
which(Rsq<0.90)

summary(fit.cont$r.squared)
```

The BALB/c sample 6 could be removed

###Determining DE genes
```{r}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```
#### DE Analysis using voom normalized counts
```{r}
results<-decideTests(fit.cont,adjust.method="none", p=0.005)
summary(results)

write.fit(fit.cont,file="IgA_Tears_DGE.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("IgA_Tears_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]
write.csv(fitObjDE,file="IgA_Tears_Voom_Hits.csv", sep=',')
fitObjDE$Autoantigen <- rownames(fitObjDE)

mydata <- as.matrix(v$E)
```

#### Boxplots from voom normalized counts ####

```{r Boxplots for Tear IgA DE Genes}
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=18, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=20), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

hits <- rownames(fitObjDE)

Y=matrix(nrow=32,ncol=15)

for (i in 1:32) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Sample <- paste0(Y$Strain, c(1:5,1:4, 1:6))

for (i in 1:32){
   filename <- paste(hits[i],"IgA_Tears.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
      scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
      ylab("Log Norm Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
   tiff(filename, units="in", width=3.45, height=3.5, res=300)
   print(p)
   dev.off()
}
```


```{r}
chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),,
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "right",
   axis.title.y = element_text(size=19, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=20), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

fitObjDE$Autoantigen[which(fitObjDE$t.NTvBT>0)]
Y_combined <- Y[,c(fitObjDE$Autoantigen[which(fitObjDE$t.NTvBT>0)])]
Y_combined[,17:18] <- Y[,33:34]
Y_combined <- Y_combined[,-c(4,12)] #Removed "GP210", "NXP2"
Y_combined <- gather(Y_combined, "Antigen", "V Counts", 1:14)

tiff("Tear_IgA_hits.tiff", units="in", width=11.5, height=8, res=300)
ggplot(Y_combined, aes(x=Strain, y=`V Counts`, fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.6, coef=1, varwidth=F, show.legend = T, size=0.7, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.2, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
   facet_wrap(~Antigen, ncol=5) +
      theme_minimal() +
      chart_design +
      ylab("Log Normalized Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
dev.off()
```



############ SERUM IgA ##################
```{r}

```

```{r}

```

