---
title: "NOD NOR Tear Serum"
author: "Shruti Kakan"
date: "2024-06-26"
output: html_document
---
#### Loading Libraries ####
```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)
library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
#library(PCAtools)
library(tidyverse)
library(Biobase)
#library(marray)
library(limma)
library(gplots)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
#devtools::install_github('renozao/NMF@devel')
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
#library(discreteMTP)
library(scales)
library(ggsci)
library(ggplot2)

```

# IgA Analysis ##########
#______________________________________________________________

```{r Reading in Raw Tear Data}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/'
)

setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/')
IgA_NSI <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_NSI_norm.csv",  header=T)[1:80,]
IgA_SNR <- read.csv("IgA_data/IgA_MCF_SSK_546_Tear_SNR.csv",  header=T)[1:80,]
Strain <- c(rep("NOD", each=5),rep("NOR", each=3), rep("BALBc", each=6))
```

#Adding Auto-antigen id names 
```{r Adding Auto-antigen id names }
#Adding Auto-antigen id names 
Antigen_ID <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/Antigen_ID.csv",  header=T)[1:80,1:2]

#colnames(IgA_NSI)[1] <- colnames(Antigen_ID)[2]
#colnames(IgA_SNR)[1] <- colnames(Antigen_ID)[2]

Antigen_ID[68,"ID"] <- IgA_NSI[68,"ID"]
IgA_NSI <- full_join(IgA_NSI, Antigen_ID, by="ID")
IgA_SNR <- full_join(IgA_SNR, Antigen_ID, by="ID")
rownames(IgA_NSI) <- IgA_NSI$Antigen_ID
rownames(IgA_SNR) <- IgA_SNR$Antigen_ID

IgA_NSI <- IgA_NSI[,-c(1, 17)]
IgA_SNR <- IgA_SNR[,-c(1, 17)]

```


```{r}
####Setting up column metadata

Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))

colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain))
colnames(colData) <- c('Sample', "Strain")
rownames(colData) <- colData$Sample
colData$Strain <- factor(colData$Strain)
#colData$Strain <- relevel(colData$Strain, ref = "BALBc")
#Biofluid <- c( rep("Tear", each=11), rep("Serum", each=11))

#colData <- as.data.frame(cbind(c(colnames(IgA_NSI)), Strain, Biofluid))

```


#### Filtering Data based on low Signal to Noise ratio
```{r Filtering rows with SNR<3}
IgA_raw=list()
IgA_SNR$average <- rowMeans(as.matrix(IgA_SNR))
IgA_SNR$med <- rowMedians(as.matrix(IgA_SNR))

IgA_raw$NSI <- as.matrix(IgA_NSI[which(IgA_SNR$average>2.8),])
IgA_raw$SNR <- as.matrix(IgA_SNR[which(IgA_SNR$average>2.8),])[,1:15]
#IgA_raw$NSI <- as.matrix(IgA_NSI[which(rowSums(IgA_SNR[,1:15]>2.8) > 6),])
#IgA_raw$SNR <- as.matrix(IgA_SNR[which(rowSums(IgA_SNR[,1:15]>2.8) > 6),][,1:15])

log2(colSums(IgA_raw$NSI*IgA_raw$SNR))

```

#### Visualizing Filtered Data ####
```{r}
dataN <- log2(IgA_NSI * IgA_SNR[,1:15] + 0.5) 
countData = as.data.frame(dataN)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count")# reshape the matrix 

mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Samples)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   #facet_wrap(~Strain, ncol=2) +
   theme_minimal() + #xlim(-1.5,6) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")
```


```{r}
boxplot(as.data.frame((dataN)),main="IgG normalization")#,col=Sample)

dataN <- (IgA_NSI * IgA_SNR[,1:15] ) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI*SNR")

dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="Log2 NSI*SNR")

dataN <- log2(IgA_raw$NSI + 0.5) 
countData = as.data.frame(dataN)
boxplot(as.data.frame((dataN)),main="NSI")

```

#Limma Based Quantile Normalization
```{r}
library(limma)
dataN <- log2(IgA_raw$NSI * IgA_raw$SNR +0.5)[,-15]
mydata <- as.matrix(dataN)

conditions<- paste(colData$Strain[-15],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)

cont.matrix<- makeContrasts(
   NTvBT =  NOD - BALBc,
   nTvBT = NOR - BALBc,
   levels = design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)

plot.new()
plotMD(fit.cont, col=2)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(fit)
```


#### Voom Normalization After Removing outliers
```{r}
#After removing outlier samples

###### Voom normalization with quantiles
conditions<- paste(colData$Strain[-15],sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

v <- voom(counts=(IgA_raw$NSI[,-15]*IgA_raw$SNR[,-15]+0.5), design, plot=TRUE, normalize="quantile")
#v <- voom(dae, design, plot=TRUE)
fit <- lmFit(v, design)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)

NTvsBT <- topTable(fit.cont, coef=1, p.value=1, number=45, adjust.method = 'BH')
nTvsBT <- topTable(fit.cont, coef=2, p.value=1, number=45, adjust.method = 'BH')
NTvsBT$Antigen <- row.names(NTvsBT)
nTvsBT$Antigen <- row.names(nTvsBT)
NTvnTvBT <- full_join(NTvsBT, nTvsBT, by="Antigen")
rownames(NTvnTvBT) <- NTvnTvBT$Antigen
NTvnTvBT[is.na(NTvnTvBT)] <- 0.5
NTvnTvBT <- NTvnTvBT[-c(which(NTvnTvBT$adj.P.Val.x > 0.05 & NTvnTvBT$adj.P.Val.y > 0.05)),]
write.csv(NTvnTvBT, file="NTvnTvBT_Tear_IgA.csv", sep=',')

#plot(x=colSums(IgA_raw$NSI*IgA_raw$SNR), y=colSums(IgG_raw$NSI*IgG_raw$SNR))
```

#### QC plots of analysis
```{r}
plotMDS(v,col=as.numeric(Strain))
plot.new()
plotMD(fit.cont, col=2)
plotMD(fit.cont, col=1)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)
```


```{r}
#Density plots
dae <- DGEList(counts=(IgA_raw$NSI*IgA_raw$SNR+0.5))
keep <- filterByExpr(dae, design)
dae <- dae[keep,,keep.lib.sizes=FALSE]
dae <- calcNormFactors(dae)
logCPM <- cpm(dae, log=TRUE, prior.count=2)

df_dseq = melt(v$E, variable.name = "Samples", value.name = "count")# reshape the matrix 
df_dseq$Strain <- factor(substr(df_dseq$Var2, 1,3))
mycolors <- colorRampPalette(brewer.pal(8,"Set1"))(15)

ggplot(df_dseq, aes(x = count, color=Var2)) + 
   geom_density(alpha = 0.5, size = 0.8) + 
   facet_wrap(~Strain, ncol=2) +
   theme_minimal() + xlim(-5, 25) +
   scale_colour_manual(values=mycolors, name="") +
   guides(fill="none")

#R-Squared ....... goodness of fit
for (i in 1:69){
   sst <- rowSums(v$E^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
plot(1:69, Rsq)
which(Rsq<0.90)

summary(fit.cont$r.squared)
```


The BALB/c sample 6 was removed

###Determining DE genes
```{r}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```
#### DE Analysis using voom normalized counts
```{r}
results<-decideTests(fit.cont,adjust.method="none", p=0.005)
summary(results)

write.fit(fit.cont,file="IgA_Tears_DGE.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("IgA_Tears_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]
write.csv(fitObjDE,file="IgA_Tears_Voom_Hits.csv", sep=',')
fitObjDE$Autoantigen <- rownames(fitObjDE)

```

#### Boxplots from voom normalized counts ####

```{r Boxplots for Tear IgA DE Genes}
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=14),
   axis.title.x = element_blank(),
   legend.text = element_text(size=15),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=18, margin = margin(r = 5)),
   strip.text.x = element_text(size =16, margin = margin(b=20), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

mydata <- as.matrix(v$E)
hits <- rownames(NTvnTvBT)
hits[(length(hits)+1)] <- "Ro SSA 52"
#hits[(length(hits)+1)] <- "RoSSA 60"

Y=matrix(nrow=40,ncol=14)

for (i in 1:40) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample[-15]

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain[-15]
Y$Sample <- paste0(Y$Strain, c(1:5,1:4, 1:5))

setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
for (i in length(hits)){
   filename <- paste(hits[i],"IgA_Tears.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.7, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9))+
      scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
      ylab("Log Norm Intensity")  + 
      scale_x_discrete(labels= c("BALB/c", "NOD", "NOR")) + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_fill_jco()
   tiff(filename, units="in", width=3.45, height=3.5, res=300)
   print(p)
   dev.off()
}
```


```{r}
chart_design <- theme(
   #plot.title = element_text(color = "Black", size = 16, face = "bold", margin = margin(b=15), hjust=0.4),,
   axis.text.x = element_text(size=16, margin = margin(b=5)),
   axis.text.y = element_text(size=18),
   axis.title.x = element_blank(),
   legend.text = element_blank(),
   legend.title = element_blank(),
   legend.position = "top",
   axis.title.y = element_text(size=24, margin = margin(r = 5)),
   strip.text.x = element_text(size =19, margin = margin(b=15), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

Y_combined <- Y[,c(NTvnTvBT$Antigen[which(NTvnTvBT$adj.P.Val.x<0.05 & NTvnTvBT$adj.P.Val.y < 0.05 & NTvnTvBT$logFC.x>0)])]
#Y_combined[,14:15] <- Y[,c("IF", "La SSB")]
Y_combined[,(ncol(Y_combined)+1):(ncol(Y_combined)+2)] <- Y[,41:42]
Y_combined <- gather(Y_combined, "Antigen", "V Counts", 1:13)
#Y_combined[which(Y_combined$Strain=="BALBc"),"Strain"] = factor("BALB/c")
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/")
tiff("Tear_IgA_hits.tiff", units="in", width=12.2, height=8.5, res=300)
ggplot(Y_combined, aes(x=Strain, y=`V Counts`, fill=Strain)) +
      geom_boxplot(outlier.shape = NA, width = 0.7, coef=1, varwidth=F, show.legend = F, size=0.9, position = position_dodge(0.9)) +
      geom_jitter(color = "darkgray", alpha =0.6, size=3.3, show.legend = F, position = position_jitterdodge(dodge.width=1.2))+
   facet_wrap(~Antigen, ncol=5, scales="free_x") +
      theme_minimal() +
      chart_design +
      ylab("Log Normalized Intensity")  + 
   scale_x_discrete(labels= c("BALB/c", "NOD", "NOR")) +
      labs(title=factor(Y_combined$Antigen), hjust=0.5) +
      scale_fill_jco()
dev.off()
```



##### Code for Comparing Tear IgG and IgA  ##################
```{r}
Tear_IgAhits <- read.csv("IgA_Tears_Voom_Hits.csv", header=T)
Tear_IgGhits <- read.csv("IgG_Tears_Voom_Hits.csv", header=T)

Tear_Ig <- full_join(Tear_IgAhits[,c(1,5,6,7,8)], Tear_IgGhits[,c(1,5,6,7,8)], by="X")

write.csv(Tear_Ig,file="IgA_IgG_Overlap_Tears_Hits.csv", sep=',')
```

#####Total tear IgG intesity values and total IgA intensity values. 
```{r}
Tear_IgG_NSI_raw <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgG_MCF_SSK_546_Tear_NSI.csv", header=TRUE)
Tear_IgG_NSI_raw <- Tear_IgG_NSI_raw[1:89,1:16]

Tear_IgA_NSI_raw <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/IgA_data/IgA_MCF_SSK_546_Tear_NSI.csv", header=TRUE, row.names = 1)
Tear_IgA_NSI_raw <- Tear_IgA_NSI_raw[1:89,]

x = colSums(Tear_IgG_NSI_raw[1:80,2:16])
y = colSums(Tear_IgA_NSI_raw[1:80,1:15])

data <- as.data.frame(cbind(x,y))
colnames(data) <- c("IgG","IgA")
Strain <- c(rep("NOD", each=5),rep("NOR", each=4), rep("BALBc", each=6))
data$Strain <- factor(Strain)
data$Sample <- rownames(data)
data <- gather(data, "Ig", "Intensity", 1:2)

tiff("ColSum_IgA_vs_IgG.tiff", units="in", width=6, height=4.5, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
ggplot(data, aes(x=Strain, y=(Intensity), fill=Ig)) +
   geom_boxplot(outlier.shape = NA, width = 0.8, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
   geom_jitter(aes(colour = Ig), alpha =0.6, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9)) +
   scale_fill_jco() + 
   theme_minimal() + ylab("Total Intensity of Autoantibodies") +
   chart_design + ylim(0,290000)
dev.off()
```

Chunk for plotting IgG vs IgA control for each of the samples
```{r}
x = unlist(Tear_IgG_NSI_raw[82,2:16])
y = unlist(Tear_IgA_NSI_raw[82,])

data <- as.data.frame(cbind(x,y))
colnames(data) <- c("IgG","IgA")
data$Strain <- factor(Strain)
data$Sample <- rownames(data)
data <- gather(data, "Ig", "Intensity", 1:2)

tiff("Tear_IgA_vs_IgG.tiff", units="in", width=6, height=4.5, res=300)
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/Tear_Auto_Validation_2022/")
ggplot(data, aes(x=Strain, y=(Intensity), fill=Ig)) +
   geom_boxplot(outlier.shape = NA, width = 0.8, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9)) +
   geom_jitter(aes(colour = Ig), alpha =0.6, size=2.5, show.legend = F, position = position_jitterdodge(dodge.width=0.9)) +
   scale_fill_jco() +
    theme_minimal() +
   ylab("Average Intensity of Ig Controls") +
   chart_design + ylim(0,100000)
dev.off()
   #+ ylim(10,12)
```

