---
title: "Serum_Autoantibody_2021"
knit: (function(input_file, encoding) {
    out_dir <- 'docs';
    rmarkdown::render(input_file,
      encoding=encoding,
      output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "ShrutiSinghKakan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/')

```


```{r Loading the libraries, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(emmeans)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
#library(PCAtools)
library(tidyverse)
library(Biobase)
#library(marray)
library(limma)
library(gplots)
```

##### 0 Dataset ######
```{r Dataset, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/')
data <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/IgG_MCF_SH_179_raw.csv",  header=T, row.names = 1)
#Columns 16-47 are empty and are therefore removed
data <- as.data.frame((data))[,-c(16:47)]

#Creating metadata including Strain and age
Strain <- c(rep(c("BALBc", "NOD"), each=6), rep(c("NOD"), each=3))
Age <- c(rep(c("8wks", "28wks"), each=3), rep(c("8wks", "24wks", "33wks"), each=3))
Group <- c(rep(c("A","B","C","D","E"), each=3))
colData0 <- as.data.frame(cbind(c('B1_8', 'B2_8', 'B3_8', 'B1_28', 'B2_28', 'B3_28', 'N1_8', 'N2_8', 'N3_8', 'N1_24', 'N2_24', 'N3_24', 'N1_33', 'N2_33', 'N3_33'), Strain, Age, Group))
colnames(colData0) <- c('Sample', "Strain", "Age", "Group")
rownames(colData0) <- colData0$Sample
colnames(data) <- colData0$Sample

colData0$Strain <- factor(colData0$Strain)
colData0$Strain <- relevel(colData0$Strain, ref = "BALBc")
colData0$Age <- factor(colData0$Age, levels = c("8wks", "24wks", "28wks", "33wks"))
colData0$Age <- relevel(colData0$Age, ref='8wks')

```

###### 1 Visualizing raw data ###########
```{r Visualization, message=FALSE, warning=FALSE, paged.print=FALSE}
#Boxplot of raw Signal Distribution, not normalized

mydata<-as.matrix(log(data/IgG_SNR[,1:15]+0.5))

boxplot(as.data.frame(mydata),main="Signal distribution, Not normalized",xlab="Samples",ylab="Log2 Intensity",cex=0.8)
```

######2 Clustering samples based on Pearson correlation ############

```{r Sample Clustering, message=FALSE, warning=FALSE, paged.print=TRUE}
matcor<-cor(mydata)
range(as.vector(matcor))

strain<-as.character(as.numeric(colData0$Strain))
age<-as.character(as.numeric(colData0$Age))

library(gplots)
heatmap.2(matcor,trace="none",col=heat.colors(40),ColSideColors=strain,
          RowSideColors=age, cexCol=1,cexRow=1,labCol="")
```

###### 3 Data filtering and Normalization ########
```{r Data filtering, message=FALSE, warning=FALSE, paged.print=FALSE}
mydata <- as.matrix((data))
dataN <- (mydata[1:95,]/mydata[96,] * 100) #df with all rows
dataN <- mydata
boxplot(as.data.frame(log2(dataN+0.5)),main="Signal distribution, Not normalized",xlab="Samples",ylab="Log2 Intensity",cex=0.8)

library(matrixStats) #for function rowmedians
IgG_SNR <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/IgG_MCF_SNR.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR$average <- rowMeans(IgG_SNR)
IgG_SNR$med <- rowMedians(as.matrix(IgG_SNR))
mydata <- data[which(IgG_SNR$average>3),] #filtering rows with high noise (i.e., rows with Signal to Noise Ratio or SNR > 3)
mydata<-as.matrix(mydata)
dataN <- mydata
boxplot(as.data.frame(log2(dataN+0.5)),main="Signal distribution, Not normalized SNR>3",xlab="Samples",ylab="Log2 Intensity",cex=0.8)

dataN <- (mydata[1:69,]/mydata[70,]) * 100000
boxplot(as.data.frame(log10(dataN+0.5)),main="IgG normalization",col=strain)

library(limma) #for function plotMDS
plotMDS(log2(dataN+0.5))
mm <- model.matrix(~0 + Group)
y <- voom((dataN), mm, plot = T)

countData = as.data.frame(y$E)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count") # reshape the matrix 
df_dseq = data.frame(df_dseq, Condition = substr(df_dseq$Samples,1,1))
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(13)

ggplot(df_dseq, aes(x = (count), colour = Samples, fill = Samples)) + 
   geom_density(alpha = 0, size = 0.8) + 
   facet_wrap(~Condition, ncol=3) #+
   #theme_minimal() + xlim(-1.5,6) +
   #scale_colour_manual(values=mycolors, name="") + guides(fill="none")
```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
matcor<-cor(dataN)
range(as.vector(matcor))
strain<-as.character(as.numeric(colData0$Strain))
age<-as.character(as.numeric(colData0$Age))
heatmap.2(matcor,trace="none",col=heat.colors(40),ColSideColors=strain,RowSideColors=age, cexCol=1,cexRow=1,labCol="")
```

###### 4 DGE analysis ##########
```{r DGE using Limma lmFit, message=FALSE, warning=FALSE, paged.print=FALSE}
#mydata <- as.matrix((data+0.5))

conditions<- paste(colData0$Strain[1:12], colData0$Age[1:12],sep=".")

conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

#contrasts of the previous analysis. 
#cont.matrix<- makeContrasts(B8vsN8 = NOD.8wks - BALBc.8wks, B28vN24 = NOD.24wks - BALBc.28wks,
#   B28vN33 = NOD.33wks - BALBc.28wks,
#   N24vN8 = NOD.24wks - NOD.8wks,
#   N33vN8 = NOD.33wks - NOD.8wks,
#   N33vN24 = NOD.33wks - NOD.24wks,
#   levels = design)
#Introducing new contrasts by removing the 33 week old samples.

cont.matrix<- makeContrasts(
B8vsN8 = NOD.8wks - BALBc.8wks,
B28vN24 = NOD.24wks - BALBc.28wks,
N24vN8 = NOD.24wks - NOD.8wks,
B28vB8 = BALBc.28wks - BALBc.8wks,
levels = design)

v <- voom(counts=((mydata[,1:12]+0.01)), design, plot=TRUE)
fit <- lmFit(v, design, robust=TRUE)
fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)


```


```{r Deciding test}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
   {
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```


```{r Limma Analysis QC plots}
plotMDS(fit.cont,col=as.numeric(Strain))
plot.new()
plotMD(fit.cont, col=2)
plotMD(fit.cont, col=1)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual)
plotDensities(v, group=Strain, col=c("orange","green", "blue"), log=TRUE)
```


```{r DGE Results using Limma, message=FALSE, warning=FALSE, paged.print=FALSE}
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/")
results<-decideTests(fit.cont,adjust.method="none", p=0.01)
summary(results)
write.fit(fit.cont,file="IgG_DGE_Limma_fit.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("IgG_DGE_Limma_fit.csv", header=T, row.names = 1)

toptable1 <- topTable(fit.cont, adjust.method="none", p=0.01, number = 30)
write.csv(toptable1,file="IgG_DGE_Limma_fitII.csv")

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]

fitObjDE$Autoantigen <- rownames(fitObjDE)
fitObjDE

```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(devtools)
install_github("dpgaile/AutoAntArrayExmpl")
#library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
#library(discreteMTP)
```

### Reading in Raw Data
```{r}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/'
)

IgG_NSI <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/IgG_MCF_NSI.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR <- read.csv("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/IgG_MCF_SNR.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR$average <- rowMeans(IgG_SNR) #adding columns for average SNR for each row (autoantibody)
IgG_SNR$med <- rowMedians(as.matrix(IgG_SNR)) #adding column for median SNR for each row 
IgG_NSI <- IgG_NSI[which(IgG_SNR$average>2),]
IgG_SNR <- IgG_SNR[which(IgG_SNR$average>2),]
IgG_raw=list()
IgG_raw$NSI <- as.matrix(IgG_NSI)
IgG_raw$SNR <- as.matrix(IgG_SNR[,1:15]) 

colData0$Strain <- relevel(colData0$Strain, ref = "BALBc")
colData0$Age <- factor(colData0$Age, levels = c("8wks", "24wks","28wks", "33wks"))
colData0$Age <- relevel(colData0$Age, ref="8wks")
IgG_raw$SInfo <- colData0

#Defining colors and point shapes
clrs=c(rep(pal_jco("default")(5), each=3))#[1:12]
pchs=c(rep(3,3),rep(4,3), rep(5,3), rep(6,3), rep(7,3))#[1:12]

```


### Sample Spot and Background Distributions
```{r}
#Centering and Scaling Differences Across Features
# Tukey Tri-mean for location
TriMnG=RowTriMeans(IgG_raw$NSI)
# Biweight midvariance for spread
bw.estG=unlist(apply(IgG_raw$NSI,1,r.bw))
```

#### Normalization First Stage ####

```{r}
## Normalization
### First Stage

#enable to background correct
IgG_raw$NSI <- as.matrix(IgG_NSI[,1:12]) #Not Excluding the Nod 33 week sample
IgG_raw$SNR <- as.matrix(IgG_SNR[,1:12])
IgG_raw$NSI=IgG_raw$NSI * IgG_raw$SNR ##NSI changed permanently
IgG_raw$X_norm_1=IgG_raw$NSI
eps=.10
IgG_raw$QregFits=array(NA,dim=c(12,70))
TriMn=RowTriMeans(IgG_raw$X_norm_1)
ResMat=IgG_raw$X_norm_1-matrix(rep(TriMn,12),ncol=12)
SubMat=abs(ResMat)<quantile(as.vector(abs(ResMat)),prob=1-eps)

par(mfrow=c(2,2));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.5,0.25))
plot(rep(TriMn,12),as.vector(ResMat),
     xlim=quantile(TriMn,prob=c(0,.8)),
     ylim=quantile(as.vector(ResMat),prob=c(0.05,.95)),
     type="n",xlab="Raw Signal Tri-Mean",ylab="Raw Signal",
     main="IgG Raw Signal")
for(j in 1:12){
   points(TriMn,ResMat[,j],cex=0.5,pch=pchs[j],col=clrs[j])
   y=ResMat[,j]
   x=TriMn
   fitj <- rq(y ~ x, tau = .5,subset=which(SubMat[,j]))
   abline(coef(fitj),col=clrs[j])
   IgG_raw$QregFits[j,]=coef(fitj)
   prdct=coef(fitj)[1]+coef(fitj)[2]*TriMn
   IgG_raw$X_norm_1[,j]=IgG_raw$X_norm_1[,j]-prdct
} ###substracting the model fit straight line from the data
#Xnorm_1 here is the residual value

eps=0.10
q.eps=0.10
lambda=.10 # incremental change
nitr=99 # number of iterations -> this should help us optimize the lambda for best fit
sad=rep(0,nitr) # sum of absolute deviations
# re-init
IgG_raw$X_1step_norm_1=IgG_raw$X_norm_1 #calculated in the last loop
IgG_raw$X_norm_1=IgG_raw$NSI #reinitializing

for(i in 1:nitr){
   # snoop and grab invariants
   X=IgG_raw$X_norm_1
   iWRS=function(i) t.test(X[i,1:6],X[i,7:12])$p.value
   qMat=matrix(rep(mapply(iWRS,1:73),12),ncol=12)
   qMat[which(is.na(qMat[,1])==T),] <- 0 #all NANs replaced by 0
   qcut=quantile(qMat[,1],prob=q.eps)
   TriMn=RowTriMeans(IgG_raw$X_norm_1)
   ResMat=IgG_raw$X_norm_1-matrix(rep(TriMn,12),ncol=12)
   SubMat=abs(ResMat)<quantile(as.vector(abs(ResMat)),prob=1-eps) #all residuals with absolute values less than the cutoff of 90%
   SubMat[qMat<qcut]=FALSE #which p values are greater than the cutoffs
   for(j in 1:12){
      y=ResMat[,j]
      x=TriMn
      fitj <- rq(y ~ x, tau = .5,subset=which(SubMat[,j]))
      prdct=coef(fitj)[1]+coef(fitj)[2]*TriMn
      IgG_raw$X_norm_1[,j]=IgG_raw$X_norm_1[,j]-lambda*prdct
      sad[i]=sad[i]+sum(abs(2*lambda*prdct))
   }###substracting the model fit straight line from the data
}
```

#### Second Stage of Normalization

```{r}
### Second Stage ########
deltaG=min(as.vector(IgG_raw$X_norm_1))
IgG_raw$W_1=log(IgG_raw$X_norm_1-deltaG+1)
# get resids..
W_1=IgG_raw$W_1
RW=RowTriMeans(W_1)
# get residual matrices
RW_MAT=matrix(rep(RW,dim(W_1)[2]),ncol=dim(W_1)[2])
W_1=W_1-RW_MAT

W=W_1
bw.est=unlist(apply(W,2,r.bw))
# now, rescale
for(j in 1:12){
   W_1[,j]=W_1[,j]/sqrt(bw.est[j])
}

# now, get new values..
IgG_raw$W_1=RW_MAT+W_1

```

## Formal Comparison of NOD.B10 and BL/10 AutoAntigen Expresion Profiles

```{r Without 33 weeks NOD}
mydata <- as.matrix(IgG_raw$W_1[,1:12])
#mydata <- as.matrix(IgG_raw$X_nrm_1)
colData <- colData0[1:12,] #removing male NOD 33 weeks sample
colData$Strain <- relevel(colData$Strain, ref = "BALBc")
colData$Age <- factor(colData$Age)

conditions<- paste(colData$Strain, colData$Age,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
rownames(design) <- colnames(mydata)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)
summary(fit)

cont.matrix<- makeContrasts(
   B8vsN8 = NOD.8wks - BALBc.8wks,
   B28vN24 = NOD.24wks - BALBc.28wks,
   BvsN = (NOD.24wks+NOD.8wks) - (BALBc.28wks+BALBc.8wks),
   Age = (NOD.24wks+BALBc.28wks) - (NOD.8wks+BALBc.8wks),
   levels = design)

fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual,pch=16,cex=0.5)
abline(0,1)
abline(0,2)
TopTable <- topTable(fit.cont,number=30,adjust="BH", p.value=0.05)

B8vsN8 <- topTable(fit.cont, coef=c("B8vsN8"), number=35, adjust="BH")
B28vsN24 <- topTable(fit.cont, coef="B28vN24", number=55, adjust="BH")
BvsN <- topTable(fit.cont, coef=c("BvsN"), number=35, adjust="BH")
Age <- topTable(fit.cont, coef=c("Age"), number=50, adjust="BH")

B8vsN8$Antigen <- row.names(B8vsN8)
B28vsN24$Antigen <- row.names(B28vsN24)
Age$Antigen <- row.names(Age)
B8vN8_B28vN24 <- full_join(B8vsN8, B28vsN24, by="Antigen") %>%
   full_join(Age, by="Antigen")
rownames(B8vN8_B28vN24) <- B8vN8_B28vN24$Antigen
#B8vN8_B28vN24[is.na(B8vN8_B28vN24)] <- 0.5
#B8vN8_B28vN24 <- B8vN8_B28vN24[which(B8vN8_B28vN24$adj.P.Val.x < 0.15 & B8vN8_B28vN24$adj.P.Val.y < 0.15),]
setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/")
write.csv(B8vN8_B28vN24, file="B8vN8_B28vN24_IgG.csv", sep=',')
write.csv(BvsN, file="AllBvsAllN.csv", sep=',')

fit.cont.exp <- getEAWP(IgG_raw$W_1)[["exprs"]]
qqnorm(fit.cont.exp, ylim=c(0,20))
qqline(fit.cont.exp)
#R-Squared ....... goodness of fit
for (i in 1:73){
   sst <- rowSums(mydata^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
print(Rsq)
#summary(lm.first)$r.squared

```


```{r With 33 weeks NOD}
mydata <- as.matrix(IgG_raw$W_1)
#mydata <- as.matrix(IgG_raw$X_nrm_1)
colData <- colData0[1:15,] #removing male NOD 33 weeks sample
colData$Strain <- relevel(colData$Strain, ref = "BALBc")
colData$Age <- factor(colData$Age)

conditions<- paste(colData$Strain, colData$Age,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
rownames(design) <- colnames(mydata)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)
summary(fit)

cont.matrix<- makeContrasts(
   B8vsN8 = NOD.8wks - BALBc.8wks,
   B28vB8 = BALBc.28wks - BALBc.8wks,
   B28vN24 = NOD.24wks - BALBc.28wks,
   N24vN8 = NOD.24wks - NOD.8wks,
   N33vN8 = NOD.33wks - NOD.8wks,
   B28vN33 = NOD.33wks - BALBc.28wks,
   Age = (NOD.24wks + NOD.33wks + BALBc.28wks )- (NOD.8wks + BALBc.8wks), 
   levels = design)

fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual,pch=16,cex=0.5)
abline(0,1)
topTable(fit.cont,number=30,adjust="BH")

#R-Squared ....... goodness of fit
for (i in 1:73){
   sst <- rowSums(mydata^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
print(Rsq)
```


### Deciding criteria for p-value correction
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```


```{r}
results <- decideTests(fit.cont,adjust.method="none", p=0.05)
summary(results)

write.fit(fit.cont,file="Serum_2021_IgG_DGE.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("Serum_2021_IgG_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]

fitObjDE$Autoantigen <- rownames(fitObjDE)

fitObjres<- fitObj[which(fitObj$P.value.B8vsN8 < 0.05 & abs(fitObj$Coef.B8vsN8) > 0), ]
fitObjres
fitObjres_old<- fitObj[which(fitObj$P.value.B28vN24 < 0.05 & abs(fitObj$Coef.B28vN24) > 0), ]
fitObjres_old

```

#### boxplot of young mice ####
```{r}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = '~/Documents/3_Parkinsons_disease/Autoantibody_Data/')

setwd('~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/')
chart_design <- theme(
   plot.title = element_text(color = "Black", size = 19, face = "bold", margin = margin(b=25), hjust=0.4),
   axis.text.x = element_text(size=15),
   axis.text.y = element_text(size=16),
   axis.title.x = element_blank(),
   legend.text = element_text(size=11, margin = margin(t=0, r=-1,l=-1)),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=17, margin = margin(r = 5)),
   strip.text.x = element_text(size =19, margin = margin(b=25), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

mydata <- as.matrix(IgG_raw$W_1)[1:73,1:12]
hits <- rownames(fitObjres)
hits[(length(hits)+1)] <- "Ro_SSA_52KDa"
hits[(length(hits)+1)] <- "Ro_SSA_60KDa"

Y=matrix(nrow=33,ncol=12)
for (i in 1:33) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Age <- colData$Age
Y$Group <- colData$Group
Y$AGE <- c(rep("Young", 3), rep("Old", 3), rep("Young", 3), rep("Old", 3))
Y$AGE <- as.factor(Y$AGE)

hits <- colnames(Y)[1:33] #change only colnames so figure headings are meaningful

#hits elements will form file names
colnames(Y)[2] <- "B2GP"
colnames(Y)[10] <- "IF"
colnames(Y)[12] <- "KU (P70/P80)"
colnames(Y)[13] <- "La/SSB"
colnames(Y)[17] <- "Mito Antigen"
colnames(Y)[21] <- "PM/Scl100"
colnames(Y)[23] <- "RPP P2"
#colnames(Y)[20] <- "RPP P0"
colnames(Y)[32] <- "Ro/SSA 52"
colnames(Y)[33] <- "Ro/SSA 60"
par(mfrow=c(2,5))
for (i in 1:33){
   setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/")
   filename <- paste(hits[i], "_SerumIgG.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i])) +
      geom_boxplot(aes(color=Strain), outlier.shape = NA, width = 0.40, coef=1, varwidth=F, show.legend = F, size=0.70, position = position_dodge(0.9) ) +
     geom_jitter(aes(fill=AGE), pch=21,size=2.5,  position=position_jitterdodge(0.40)) +
      scale_color_jco() +
     scale_fill_aaas() +
      guides(color="none") +
      theme_minimal() +
      chart_design +
         ylab("Log Norm Intensity")  + ylim(5.5,24.5) +
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_x_discrete("Strain", labels=c("BALB/c", "NOD"))
   tiff(filename, units="in", width=2.35, height=3.6, res=300)
   print(p)
   dev.off()
}

```

#### boxplot of old mice######
```{r}
mydata <- as.matrix(IgG_raw$W_1)[1:73,1:12]
hits <- rownames(fitObjres_old)
Y=matrix(nrow=26,ncol=12)
for (i in 1:26) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Age <- colData$Age
Y$Group <- colData$Group
Y$AGE <- c(rep("Young", 3), rep("Old", 3), rep("Young", 3), rep("Old", 3))
Y$AGE <- as.factor(Y$AGE)

#hits elements will form file names

colnames(Y)[10] <- "KU (P70/P80)"
colnames(Y)[11] <- "La/SSB"
colnames(Y)[19] <- "PM/Scl100"
colnames(Y)[21] <- "RPP P2"
colnames(Y)[22] <- "RPP P0"

#Y=Y[c(4:6, 10:12),]
par(mfrow=c(2,5))
for (i in 1:26){
   setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/")
   filename <- paste(hits[i], "IgG.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i])) +
      geom_boxplot(aes(color=Strain), outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = F, size=0.5) +
     geom_jitter(aes(fill=AGE), size=2.5, pch=21, width = 0.25) +
      scale_color_jco() +
     scale_fill_aaas() +
      theme_minimal() +
      chart_design +
         ylab("Log2 Norm Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_x_discrete("Strain", labels=c("BALB/c", "NOD")) + 
     ylim(5.5,24.5) + guides(color="none")
   tiff(filename, units="in", width=2.35, height=3.6, res=300)
   print(p)
   dev.off()
}
```


#### Boxplots of Age ###

```{r}

fitObjAge <- fitObj[which(fitObj$P.value.Age < 0.05 ), ]
mydata <- as.matrix(IgG_raw$W_1)[1:73,1:12]
hits <- rownames(fitObjAge)
Y=matrix(nrow=5,ncol=12)
for (i in 1:5) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Age <- colData$Age
Y$Group <- colData$Group
Y$AGE <- c(rep("Young", 3), rep("Old", 3), rep("Young", 3), rep("Old", 3))
Y$AGE <- as.factor(Y$AGE)
Y$AGE <- relevel(Y$AGE, ref="Young")
#hits elements will form file names

colnames(Y)[2] <- "Gly Albumin"
colnames(Y)[3] <- "PM/Scl100"
colnames(Y)[4] <- "Ro/SSA 52"

#Y=Y[c(4:6, 10:12),]
par(mfrow=c(2,5))
for (i in 1:5){
   setwd("~/Documents/3_Parkinsons_disease/Autoantibody_Data/SerumAntibodies_2021/")
   filename <- paste(hits[i], "Age.IgG.tiff", sep="")
   p<- ggplot(Y, aes(x=AGE, y=Y[,i])) +
      geom_boxplot(aes(color=AGE), outlier.shape = NA, width = 0.5, coef=1, varwidth=F, show.legend = F, size=0.5) +
     geom_jitter(aes(fill=Strain), size=2.5, pch=21, width = 0.25) +
      scale_fill_jco() +
     scale_color_aaas() +
      theme_minimal() +
      chart_design +
         ylab("Log2 Norm Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      #scale_x_discrete("Strain", labels=c("BALB/c", "NOD")) + 
     #ylim(1.5,24) + 
      guides(color="none")
   tiff(filename, units="in", width=2.35, height=3.9, res=300)
   print(p)
   dev.off()
}
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
deltaG=min(as.vector(IgG_raw$X_norm_1))
IgG_raw$W_1=log(IgG_raw$X_norm_1-deltaG+1)
IgG_raw$W_2=log(IgG_raw$X_norm_1-deltaG+1)

#tiff("fig7.tiff", units="in", width=8, height=7, res=300)  
par(mfrow=c(2,2));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.5,0.25))
AAModesIgG=AAplotDens(IgG_raw,SampleIndices=1:12,clrs=clrs,useMat="resid",xlims=NA,returnModes=T,main="IgG Residual (Stage 1) log Spot Signal")
legend(-0.6,14,lty=1,lwd=2,col=clrs[c(1,4,7,10)],legend=c(paste(rep("B",2),rep(c(8,28), each=1)), paste(rep("N",2),rep(c(8,24), each=1))),bty="n",ncol=1) #did not plot

# get resids..
W_1=IgG_raw$W_1
RW=RowTriMeans(W_1)
# get residual matrices
RW_MAT=matrix(rep(RW,dim(W_1)[2]),ncol=dim(W_1)[2])
W_1=W_1-RW_MAT #value - trimean

W=W_1
bw.est=unlist(apply(W,2,r.bw)) ####bivariate midvariance is something like pearsons r2. variances across columns

# now, rescale
for(j in 1:12){
   W_1[,j]=W_1[,j]/sqrt(bw.est[j]) #sq. root of bivariance is an estimate of sd
}
# now, get new values..
IgG_raw$W_1=RW_MAT+W_1
IgG_raw$W_2=RW_MAT+W_1

AAplotDens(IgG_raw,SampleIndices=1:12,clrs=clrs,useMat='resid',
           xlims=NA,returnModes=T,main="IgG Residual (Stage 2) log Spot Signal")
legend(2.5,0.55,lty=1,lwd=2,col=clrs,legend=c(paste(rep("B",2),rep(c(8,28), each=1)), paste(rep("N",2),rep(c(8,24), each=1))),bty="n",ncol=1)

plot(density(log2(IgG_raw$W_1[,1]+65)-7.20), main="(logW_1) ", sub=" . ", type="l")
for(j in 2:12) {
   dens<-density(log2(IgG_raw$W_1[,j]+65)-7.20)
   lines(dens, cex=0.5,pch=pchs[j],col=clrs[j],cex.main=1)
}

AAplotByMean(IgG_raw,SampleIndices=1:12,clrs=clrs,pchs=pchs,MeanSubset=1:12,useMat="log",main="IgG log normalized")
AAplotByMean(IgG_raw,SampleIndices=1:12,clrs=clrs,pchs=pchs,MeanSubset=1:12,useMat="resid",main="IgG residuals log normalized")
#dev.off()

```

###### Heatmap of normalized data
```{r Heatmap final, eval=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
###### Heatmap of normalized data

RnkSmplW_1=matrix(nrow=70,ncol=12)
RnkSmplW_1[,1]=IgG_raw$W_1[,1]
for(j in 2:12) RnkSmplW_1[,j]=IgG_raw$W_1[,j]
for(i in 1:70) RnkSmplW_1[i,]=rank(RnkSmplW_1[i,])/(13)
colnames(RnkSmplW_1)=as.character((colData$Sample))
rownames(RnkSmplW_1) <- rownames(IgG_raw$W_1)

#tiff("fig8.tiff", units="in", width=6, height=12.5, res=300)   
pheatmap(RnkSmplW_1,annCol=factor(colData$Strain),main="Rank-it (Across Samples) Normalized Signal IgG",dist="euclidean",hclust="ward")
#dev.off()
```


```{r}

```


```{r}

```



```{r}

```


```{r}

```

