---
title: "Serum_Autoantibody_2021"
knit: (function(input_file, encoding) {
    out_dir <- 'docs';
    rmarkdown::render(input_file,
      encoding=encoding,
      output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "ShrutiSinghKakan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Loading the libraries, message=FALSE, warning=FALSE}

library(car)
library(lsmeans)
library(calibrate)
library(dplyr)
#library(DEGreport)
library(DESeq2)
library(DEFormats)
library(edgeR)
library(emmeans)
library(ggpubr)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(pheatmap)
library(reshape2)
library(RColorBrewer)
library(scales)
library(rstatix)
library(tidyr)
library(magrittr)
library(PCAtools)
library(tidyverse)
library(Biobase)
library(marray)
library(limma)
library(gplots)
```

##### 0 Dataset ######
```{r Dataset, message=FALSE, warning=FALSE, paged.print=FALSE}
data <- read.csv("../../SerumAntibodies_2021/IgG_MCF_SH_179_raw.csv",  header=T, row.names = 1)
#Columns 16-47 are empty and are therefore removed
data <- as.data.frame((data))[,-c(16:47)]

#Creating metadata including Strain and age
Strain <- c(rep(c("BALBc", "NOD"), each=6), rep(c("NOD"), each=3))
Age <- c(rep(c("8wks", "28wks"), each=3), rep(c("8wks", "24wks", "33wks"), each=3))
Group <- c(rep(c("A","B","C","D","E"), each=3))
colData <- as.data.frame(cbind(c('B1_8', 'B2_8', 'B3_8', 'B1_28', 'B2_28', 'B3_28', 'N1_8', 'N2_8', 'N3_8', 'N1_24', 'N2_24', 'N3_24', 'N1_33', 'N2_33', 'N3_33'), Strain, Age, Group))
colnames(colData) <- c('Sample', "Strain", "Age", "Group")
rownames(colData) <- colData$Sample
colnames(data) <- colData$Sample

colData$Strain <- factor(colData$Strain)
colData$Strain <- relevel(colData$Strain, ref = "BALBc")
colData$Age <- factor(colData$Age, levels = c("8wks", "24wks", "28wks", "33wks"))
colData$Age <- relevel(colData$Age, ref='8wks')

```

###### 1 Visualizing raw data ###########
```{r Visualization, message=FALSE, warning=FALSE, paged.print=FALSE}
#Boxplot of raw Signal Distribution, not normalized

mydata<-as.matrix(log2(data+0.5))

boxplot(as.data.frame(mydata),main="Signal distribution, Not normalized",xlab="Samples",ylab="Log2 Intensity",cex=0.8)
```

######2 Clustering samples based on Pearson correlation ############

```{r Sample Clustering, message=FALSE, warning=FALSE, paged.print=TRUE}
matcor<-cor(mydata)
range(as.vector(matcor))

strain<-as.character(as.numeric(colData$Strain))
age<-as.character(as.numeric(colData$Age))

library(gplots)
heatmap.2(matcor,trace="none",col=heat.colors(40),ColSideColors=strain,
          RowSideColors=age, cexCol=1,cexRow=1,labCol="")
```

###### 3 Data filtering and Normalization ########
```{r Data filtering, message=FALSE, warning=FALSE, paged.print=FALSE}
mydata <- as.matrix((data))
dataN <- (mydata[1:95,]/mydata[96,] * 100) #df with all rows
dataN <- mydata
boxplot(as.data.frame(log2(dataN+0.5)),main="Signal distribution, Not normalized",xlab="Samples",ylab="Log2 Intensity",cex=0.8)

library(matrixStats) #for function rowmedians
IgG_SNR <- read.csv("../../SerumAntibodies_2021/IgG_MCF_SNR.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR$average <- rowMeans(IgG_SNR)
IgG_SNR$med <- rowMedians(as.matrix(IgG_SNR))
mydata <- data[which(IgG_SNR$average>3),] #filtering rows with high noise (i.e., rows with Signal to Noise Ratio or SNR > 3)
mydata<-as.matrix(mydata)
dataN <- mydata
boxplot(as.data.frame(log2(dataN+0.5)),main="Signal distribution, Not normalized SNR>3",xlab="Samples",ylab="Log2 Intensity",cex=0.8)

dataN <- (mydata[1:69,]/mydata[70,]) * 100000
boxplot(as.data.frame(log10(dataN+0.5)),main="IgG normalization",col=strain)

library(limma) #for function plotMDS
plotMDS(log2(dataN+0.5))
mm <- model.matrix(~0 + Group)
y <- voom((dataN), mm, plot = T)

countData = as.data.frame(dataN)

df_dseq = melt(countData, variable.name = "Samples", value.name = "count") # reshape the matrix 
df_dseq = data.frame(df_dseq, Condition = substr(df_dseq$Samples,1,1))
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(13)

ggplot(df_dseq, aes(x = log10(count+0.5), colour = Samples, fill = Samples)) + 
   geom_density(alpha = 0, size = 0.8) + 
   facet_wrap(~Condition, ncol=3) #+
   #theme_minimal() + xlim(-1.5,6) +
   #scale_colour_manual(values=mycolors, name="") + guides(fill="none")
```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
matcor<-cor(dataN)
range(as.vector(matcor))
strain<-as.character(as.numeric(colData$Strain))
age<-as.character(as.numeric(colData$Age))
heatmap.2(matcor,trace="none",col=heat.colors(40),ColSideColors=strain,RowSideColors=age, cexCol=1,cexRow=1,labCol="")
```

###### 4 DGE analysis ##########
```{r DGE using Limma lmFit, message=FALSE, warning=FALSE, paged.print=FALSE}
#mydata <- as.matrix((data+0.5))

conditions<- paste(colData$Strain[1:12], colData$Age[1:12],sep=".")

conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
colnames(design) <- levels(conditions)

#contrasts of the previous analysis. 
#cont.matrix<- makeContrasts(B8vsN8 = NOD.8wks - BALBc.8wks, B28vN24 = NOD.24wks - BALBc.28wks,
#   B28vN33 = NOD.33wks - BALBc.28wks,
#   N24vN8 = NOD.24wks - NOD.8wks,
#   N33vN8 = NOD.33wks - NOD.8wks,
#   N33vN24 = NOD.33wks - NOD.24wks,
#   levels = design)
#Introducing new contrasts by removing the 33 week old samples.
fit <- lmFit(dataN[,1:12], design)
cont.matrix<- makeContrasts(
B8vsN8 = NOD.8wks - BALBc.8wks,
B28vN24 = NOD.24wks - BALBc.28wks,
N24vN8 = NOD.24wks - NOD.8wks,
B28vB8 = BALBc.28wks - BALBc.8wks,
levels = design)

fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)

decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
   {
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```


```{r DGE Results using Limma, message=FALSE, warning=FALSE, paged.print=FALSE}
results<-decideTests(fit.cont,adjust.method="none", p=0.01)
summary(results)
write.fit(fit.cont,file="IgG_DGE_0423.csv",adjust="none",results=results, sep=',')
fitObj <- read.csv("IgG_DGE_0423.csv", header=T, row.names = 1)

toptable1 <- topTable(fit.cont, coef=2, adjust.method="none", p=0.01, sort.by = "P", n = Inf)
write.csv(toptable1,file="IgG_DGE_0423_II.csv")

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]

fitObjDE$Autoantigen <- rownames(fitObjDE)
fitObjDE

```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)

library(devtools)
#install_github("dpgaile/AutoAntArrayExmpl")
library(AutoAntArrayExmpl)
library(NMF)
library(quantreg)
library(asbio)
library(fdrtool)
library(discreteMTP)
```

### Reading in Raw Data
```{r}
IgG_NSI <- read.csv("../../SerumAntibodies_2021/IgG_MCF_NSI.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR <- read.csv("../../SerumAntibodies_2021/IgG_MCF_SNR.csv",  header=T, row.names = 1)[,1:15]
IgG_SNR$average <- rowMeans(IgG_SNR) #adding columns for average SNR for each row (autoantibody)
IgG_SNR$med <- rowMedians(as.matrix(IgG_SNR)) #adding column for median SNR for each row 
IgG_NSI <- IgG_NSI[which(IgG_SNR$average>3),]
IgG_SNR <- IgG_SNR[which(IgG_SNR$average>3),]
IgG_raw=list()
IgG_raw$NSI <- as.matrix(IgG_NSI[,1:12])
IgG_raw$SNR <- as.matrix(IgG_SNR[,1:12]) #Excluding the Nod 33 week sample

colData$Strain <- relevel(colData$Strain, ref = "BALBc")
colData$Age <- factor(colData$Age, levels = c("8wks", "24wks","28wks"))
colData$Age <- relevel(colData$Age, ref="8wks")
IgG_raw$SInfo <- colData


#Defining colors and point shapes
clrs=c(rep(pal_jco("default")(5), each=3))#[1:12]
pchs=c(rep(3,3),rep(4,3), rep(5,3), rep(6,3), rep(7,3))#[1:12]

```


### Sample Spot and Background Distributions
```{r}
#Centering and Scaling Differences Across Features
# Tukey Tri-mean for location
TriMnG=RowTriMeans(IgG_raw$NSI)
# Biweight midvariance for spread
bw.estG=unlist(apply(IgG_raw$NSI,1,r.bw))
```

#### Normalization First Stage ####

```{r}
## Normalization
### First Stage

#enable to background correct
IgG_raw$NSI <- as.matrix(IgG_NSI[,1:12])
IgG_raw$NSI=IgG_raw$NSI * IgG_raw$SNR ##NSI changed permanently
IgG_raw$X_norm_1=IgG_raw$NSI
eps=.10
IgG_raw$QregFits=array(NA,dim=c(12,70))
TriMn=RowTriMeans(IgG_raw$X_norm_1)
ResMat=IgG_raw$X_norm_1-matrix(rep(TriMn,12),ncol=12)
SubMat=abs(ResMat)<quantile(as.vector(abs(ResMat)),prob=1-eps)

par(mfrow=c(2,2));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.5,0.25))
plot(rep(TriMn,12),as.vector(ResMat),
     xlim=quantile(TriMn,prob=c(0,.8)),
     ylim=quantile(as.vector(ResMat),prob=c(0.05,.95)),
     type="n",xlab="Raw Signal Tri-Mean",ylab="Raw Signal",
     main="IgG Raw Signal")
for(j in 1:12){
   points(TriMn,ResMat[,j],cex=0.5,pch=pchs[j],col=clrs[j])
   y=ResMat[,j]
   x=TriMn
   fitj <- rq(y ~ x, tau = .5,subset=which(SubMat[,j]))
   abline(coef(fitj),col=clrs[j])
   IgG_raw$QregFits[j,]=coef(fitj)
   prdct=coef(fitj)[1]+coef(fitj)[2]*TriMn
   IgG_raw$X_norm_1[,j]=IgG_raw$X_norm_1[,j]-prdct
} ###substracting the model fit straight line from the data
#Xnorm_1 here is the residual value

eps=0.10
q.eps=0.10
lambda=.10 # incremental change
nitr=75 # number of iterations -> this should help us optimize the lambda for best fit
sad=rep(0,nitr) # sum of absolute deviations
# re-init
IgG_raw$X_1step_norm_1=IgG_raw$X_norm_1 #calculated in the last loop
IgG_raw$X_norm_1=IgG_raw$NSI #reinitializing

for(i in 1:nitr){
   # snoop and grab invariants
   X=IgG_raw$X_norm_1
   iWRS=function(i) t.test(X[i,1:6],X[i,7:12])$p.value
   qMat=matrix(rep(mapply(iWRS,1:70),12),ncol=12)
   qMat[which(is.na(qMat[,1])==T),] <- 0 #all NANs replaced by 0
   qcut=quantile(qMat[,1],prob=q.eps)
   TriMn=RowTriMeans(IgG_raw$X_norm_1)
   ResMat=IgG_raw$X_norm_1-matrix(rep(TriMn,12),ncol=12)
   SubMat=abs(ResMat)<quantile(as.vector(abs(ResMat)),prob=1-eps) #all residuals with absolute values less than the cutoff of 90%
   SubMat[qMat<qcut]=FALSE #which p values are greater than the cutoffs
   for(j in 1:12){
      y=ResMat[,j]
      x=TriMn
      fitj <- rq(y ~ x, tau = .5,subset=which(SubMat[,j]))
      prdct=coef(fitj)[1]+coef(fitj)[2]*TriMn
      IgG_raw$X_norm_1[,j]=IgG_raw$X_norm_1[,j]-lambda*prdct
      sad[i]=sad[i]+sum(abs(2*lambda*prdct))
   }###substracting the model fit straight line from the data
}
```

#### Second Stage of Normalization

```{r}
### Second Stage ########
deltaG=min(as.vector(IgG_raw$X_norm_1))
IgG_raw$W_1=log(IgG_raw$X_norm_1-deltaG+1)
# get resids..
W_1=IgG_raw$W_1
RW=RowTriMeans(W_1)
# get residual matrices
RW_MAT=matrix(rep(RW,dim(W_1)[2]),ncol=dim(W_1)[2])
W_1=W_1-RW_MAT

W=W_1
bw.est=unlist(apply(W,2,r.bw))
# now, rescale
for(j in 1:12){
   W_1[,j]=W_1[,j]/sqrt(bw.est[j])
}

# now, get new values..
IgG_raw$W_1=RW_MAT+W_1

```

## Formal Comparison of NOD.B10 and BL/10 AutoAntigen Expresion Profiles

```{r}
mydata <- as.matrix(IgG_raw$W_1)
#mydata <- as.matrix(IgG_raw$X_nrm_1)
colData <- colData[1:12,] #removing male NOD 33 weeks sample
colData$Strain <- relevel(colData$Strain, ref = "BALBc")
colData$Age <- factor(colData$Age)

conditions<- paste(colData$Strain, colData$Age,sep=".")
conditions <- factor(conditions, levels=unique(conditions))
design <- model.matrix(~0+ conditions)
rownames(design) <- colnames(mydata)
colnames(design) <- levels(conditions)
fit <- lmFit(mydata, design)
summary(fit)

cont.matrix<- makeContrasts(
   B8vsN8 = NOD.8wks - BALBc.8wks,
   B28vB8 = BALBc.28wks - BALBc.8wks,
   B28vN24 = NOD.24wks - BALBc.28wks,
   N24vN8 = NOD.24wks - NOD.8wks,
   Age = (NOD.24wks + BALBc.28wks)- (NOD.8wks+BALBc.8wks), 
   levels = design)

fit.cont<- contrasts.fit(fit, cont.matrix)
fit.cont<- eBayes(fit.cont)
qqt(fit.cont$t,df=fit.cont$df.prior+fit.cont$df.residual,pch=16,cex=0.5)
abline(0,1)
topTable(fit.cont,number=15,adjust="BH")

#R-Squared ....... goodness of fit
for (i in 1:70){
   sst <- rowSums(mydata^2)
   ssr <- sst - fit.cont$df.residual*(fit.cont$sigma^2)
   Rsq<- (ssr/sst)
}
#summary(lm.first)$r.squared

```


### Deciding criteria for p-value correction
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
decide <- matrix(c("fdr",0.05,
                   "fdr",0.1, "none",0.005, "none", 0.01),nrow=4,ncol=2,byr=T)
# initialize:
mysum <- as.list(1:nrow(decide))
mynum <- 0
maxmax <- 0
for (test in 1:nrow(decide)){
   results<-decideTests(fit.cont,
                        adjust.method=decide[test,1],
                        p=as.numeric(decide[test,2]))
   summary(results) -> mysum[[test]]
   mynum[test] <-length(which(apply(results,1,function(x)any(x,na.rm=T))))
   maxmax <- max(c(maxmax, as.vector(mysum[[test]][c(1,3),])))
}
par(mfrow=c(1,nrow(decide)))
for (test in 1:nrow(decide))
{
   as.numeric(as.vector(mysum[[test]][3,]))->plotMe1
   as.numeric(as.vector(mysum[[test]][1,]))->plotMe2
   maxData = max(plotMe1)
   maxData2 = max(plotMe2)
   barplot(plotMe1,horiz=T,col="red",xlim=c(-maxmax,maxmax),
           main=paste("Gene Changes \np<",decide[test,2], ", " , decide[test,1],
                      " (" ,mynum[test] ,")",sep=""))->yy
   barplot(-plotMe2,horiz=T,col="green",add=T)->yy
   xx<-vector("integer",ncol(mysum[[test]]))
   text(xx,yy,colnames(mysum[[test]]))
   text((plotMe1+10)*0 + .9*maxData,yy+0.1,format(plotMe1,digits=3))
   text((-plotMe2-10)*0 - .9*maxData2,yy+0.1,format(plotMe2,digits=3))
}
```

```{r}
results <- decideTests(fit.cont,adjust.method="fdr", p=0.05)
summary(results)

write.fit(fit.cont,file="Serum_2021_IgG_DGE.csv",adjust="fdr",results=results, sep=',')
fitObj <- read.csv("Serum_2021_IgG_DGE.csv", header=T, row.names = 1)

myNames<-names(fitObj)
res.col<- which(regexpr("Res.",myNames)>0)
DElist<- which(apply(fitObj[,res.col],1,function(x)any(x,na.rm=T)))
length(DElist)

fitObjDE <-fitObj[DElist,]

fitObjDE$Autoantigen <- rownames(fitObjDE)

fitObjres<- fitObj[which(fitObj$P.value.B28vN24 < 0.05 & abs(fitObj$Coef.B28vN24) > 0), ]
fitObjres

```

#### boxplot ######
```{r}
mydata <- as.matrix(IgG_raw$W_1)[1:70,]
hits <- rownames(fitObjres)

chart_design <- theme(
   plot.title = element_text(color = "Black", size = 17, face = "bold", margin = margin(b=25), hjust=0.4),
   axis.text.x = element_text(size=14),
   axis.text.y = element_text(size=12),
   axis.title.x = element_blank(),
   legend.text = element_text(size=12, margin = margin(r=0,l=-5)),
   legend.title = element_blank(),
   legend.position = "bottom",
   axis.title.y = element_text(size=13, margin = margin(r = 5)),
   strip.text.x = element_text(size =17, margin = margin(b=25), face='bold', hjust=0.4),
   strip.background = element_blank(), 
   strip.placement = "outside")

Y=matrix(nrow=26,ncol=12)
for (i in 1:26) {
   Y[i,] <- mydata[hits[i],]
}
rownames(Y) <- hits
colnames(Y) <- colData$Sample

Y <- as.data.frame(t(Y))
Y$Strain <- colData$Strain
Y$Age <- colData$Age
Y$Group <- colData$Group

hits <- colnames(Y)[1:13]
colnames(Y)[8] <- "KU (P70/P80)"
colnames(Y)[9] <- "La/SSB"
colnames(Y)[17] <- "PM/Scl-100"
colnames(Y)[19] <- "RPP P2"
colnames(Y)[20] <- "RPP P0"

par(mfrow=c(2,5))
for (i in 1:26){
   filename <- paste(hits[i], ".IgG.tiff", sep="")
   p<- ggplot(Y, aes(x=Strain, y=Y[,i], fill=Age)) +
      geom_boxplot(outlier.shape = NA, width = 0.7, coef=1, varwidth=F, show.legend = T, size=0.9, position = position_dodge(0.9) ) +
      scale_fill_jco() +
      scale_color_manual(values=c("black", "navy")) +
      theme_minimal() +
      chart_design +
         ylab("Log Normalized Intensity")  + 
      labs(title=colnames(Y[i]), hjust=0.5) +
      scale_x_discrete("Strain", labels=c("BALB/c", "NOD"))
   tiff(filename, units="in", width=2.85, height=2.9, res=300)
   print(p)
   dev.off()
}
```


```{r}

```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
deltaG=min(as.vector(IgG_raw$X_norm_1))
IgG_raw$W_1=log(IgG_raw$X_norm_1-deltaG+1)
IgG_raw$W_2=log(IgG_raw$X_norm_1-deltaG+1)

#tiff("fig7.tiff", units="in", width=8, height=7, res=300)  
par(mfrow=c(2,2));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.5,0.25))
AAModesIgG=AAplotDens(IgG_raw,SampleIndices=1:12,clrs=clrs,useMat="resid",xlims=NA,returnModes=T,main="IgG Residual (Stage 1) log Spot Signal")
legend(-0.6,14,lty=1,lwd=2,col=clrs[c(1,4,7,10)],legend=c(paste(rep("B",2),rep(c(8,28), each=1)), paste(rep("N",2),rep(c(8,24), each=1))),bty="n",ncol=1) #did not plot

# get resids..
W_1=IgG_raw$W_1
RW=RowTriMeans(W_1)
# get residual matrices
RW_MAT=matrix(rep(RW,dim(W_1)[2]),ncol=dim(W_1)[2])
W_1=W_1-RW_MAT #value - trimean

W=W_1
bw.est=unlist(apply(W,2,r.bw)) ####bivariate midvariance is something like pearsons r2. variances across columns

# now, rescale
for(j in 1:12){
   W_1[,j]=W_1[,j]/sqrt(bw.est[j]) #sq. root of bivariance is an estimate of sd
}
# now, get new values..
IgG_raw$W_1=RW_MAT+W_1
IgG_raw$W_2=RW_MAT+W_1

AAplotDens(IgG_raw,SampleIndices=1:12,clrs=clrs,useMat='resid',
           xlims=NA,returnModes=T,main="IgG Residual (Stage 2) log Spot Signal")
legend(2.5,0.55,lty=1,lwd=2,col=clrs,legend=c(paste(rep("B",2),rep(c(8,28), each=1)), paste(rep("N",2),rep(c(8,24), each=1))),bty="n",ncol=1)

plot(density(log2(IgG_raw$W_1[,1]+65)-7.20), main="(logW_1) ", sub=" . ", type="l")
for(j in 2:12) {
   dens<-density(log2(IgG_raw$W_1[,j]+65)-7.20)
   lines(dens, cex=0.5,pch=pchs[j],col=clrs[j],cex.main=1)
}

AAplotByMean(IgG_raw,SampleIndices=1:12,clrs=clrs,pchs=pchs,MeanSubset=1:12,useMat="log",main="IgG log normalized")
AAplotByMean(IgG_raw,SampleIndices=1:12,clrs=clrs,pchs=pchs,MeanSubset=1:12,useMat="resid",main="IgG residuals log normalized")
#dev.off()

```

###### Heatmap of normalized data
```{r Heatmap final, eval=FALSE, fig.height=12, fig.width=7, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
###### Heatmap of normalized data

RnkSmplW_1=matrix(nrow=70,ncol=12)
RnkSmplW_1[,1]=IgG_raw$W_1[,1]
for(j in 2:12) RnkSmplW_1[,j]=IgG_raw$W_1[,j]
for(i in 1:70) RnkSmplW_1[i,]=rank(RnkSmplW_1[i,])/(13)
colnames(RnkSmplW_1)=as.character((colData$Sample))
rownames(RnkSmplW_1) <- rownames(IgG_raw$W_1)

#tiff("fig8.tiff", units="in", width=6, height=12.5, res=300)   
pheatmap(RnkSmplW_1,annCol=factor(colData$Strain),main="Rank-it (Across Samples) Normalized Signal IgG",dist="euclidean",hclust="ward")
#dev.off()
```

```{r Scaling, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
par(mfrow=c(1,1));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.75,1.5))
# get resids..
W_1G=IgG_raw$W_1
RWG=RowTriMeans(W_1G)
# get residual matrices
RWG_MAT=matrix(rep(RWG,dim(W_1G)[2]),ncol=dim(W_1G)[2])
res_1G=W_1G-RWG_MAT

# diagnostic.. check for comp hybrid effects..
smoothScatter(as.vector(rbind(res_1G,res_1G)))  

ResG=res_1G[,1]
for(j in 2:12){
   ResG=cbind(ResG,res_1G[,j])
}
spD_G=as.dist(1-cor(ResG,method="spearman"))

cmdG=cmdscale(spD_G,k=2)

rownames(cmdG)=as.character(c(paste(rep("B",6),rep(c(8,28), each=3)), paste(rep("N",6),rep(c(8,24), each=3))))

plot(cmdG[,1],cmdG[,2],type="n",axes=F,xlab="",ylab="",
     main="Metric Multidimensional Scaling: IgG",cex.main=1)
axis(1);axis(2)
for(i in 1:6){
   i1=2*i-1 ; i2=2*i
   lines(cmdG[c(i1,i2),1],cmdG[c(i1,i2),2],col=clrs[i])
   text(cmdG[i1,1],cmdG[i1,2],col=clrs[i],rownames(cmdG)[i1],cex=3/4)
   text(cmdG[i2,1],cmdG[i2,2],col=clrs[i],rownames(cmdG)[i2],cex=3/4)
}
```


######## Differential Expression 

#Wilcoxon Rank Sum tests were utilized to formally test for differential autoantigen expression (with respect to NOD.B10 versus BL/10 animals). P-values were adjusted by the method of Benjamini and Hochberg [@bh1995] and the corresponding "q-values" are reported. Given the discrete nature of the underlying p-value distribution, the Benjamini and Hochberg FDR estimates can be considered as conservative.  The more liberal mid-P [@heller2011false] modified Benjamini and Hochberg q-values were also calculated and are reported (although, the control of FDR for that approach is not strictly guaranteed).

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
par(mfrow=c(2,1));par(mgp=c(1.5,.5,0));par(mar=c(2.75,2.75,1.75,0.25))

#====================================================#
# Code for IgG
#====================================================#

W=IgG_raw$W_1
RTM=rank(RowTriMeans(W_1))/13

#### all balbs vs all NODs ######

#packageurl <- "http://cran.nexr.com/src/contrib/discreteMTP_0.1-2.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")
library(discreteMTP) #function p.discrete.adjust uses this package
iWRS=function(i) wilcox.test(W[i,1:6],W[i,7:12])$p.value
myWRS=mapply(iWRS,1:70)
myQ=p.adjust(myWRS,method="BH")
pCDFlist=list(2*((1:151)/252))
for(i in 2:70) pCDFlist=c(pCDFlist,list(pCDFlist[[1]]))
myQ.disc=p.discrete.adjust(myWRS, pCDFlist, method = "BHmidp")

# order the features by statistical significance
odx=order(myWRS)
# compute HC scores
hcG=hc.score(myWRS)
# determine HC threshold
hcGcut=hc.thresh(myWRS)
title("Higher Critcism Decision Threshold: IgG")
hcDX=which(myWRS[odx]<=hcGcut)
out=cbind(myWRS,myQ,myQ.disc,hcG,RTM)
rownames(out)=rownames(W)
colnames(out)=c("p-value","q-value (BH)","q-values (BH midP)","HC score","Tri-Mean rank-it")
out=out[odx,]
#cat("\n AutoAntigen IgG Differential Expression",fill=T)
#print(out[hcDX,])
DE.IgGi=which(myWRS<0.025)
# make table
tblGi=round(out[hcDX,],4)

```

```{r}

```



```{r}

```


```{r}

```



```{r}

```


```{r}

```

